---
title: "Step5:ALE"
author: "Yeonui+Kwak"
date: "8/9/2020"
output: html_document
---

1.Set a working directory
```{R}
setwd("~/Desktop/Publication_Mar/Final_Fig3/Rcode/")

```
#set up color
```{r}
library(RColorBrewer)
#nstall.packages("raster")
library(raster)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }
```

```{R}
if(!require(coin)){install.packages("coin")}
if(!require(rcompanion)){install.packages("rcompanion")}
library(coin);library(rcompanion)
#install.packages("DescTools")
#library(DescTools)
pears.cor=function(table, rscore, cscore)
{ 
	dim=dim(table) 
	rbar=sum(margin.table(table,1)*rscore)/sum(table) 
	rdif=rscore-rbar 
	cbar=sum(margin.table(table,2)*cscore)/sum(table) 
	cdif=cscore-cbar 
	ssr=sum(margin.table(table,1)*(rdif^2)) 
	ssc=sum(margin.table(table,2)*(cdif^2)) 
	ssrc=sum(t(table*rdif)*cdif) 
	pcor=ssrc/(sqrt(ssr*ssc)) 
	pcor 
	M2=(sum(table)-1)*pcor^2
	M2
	result=c(pcor, M2)
	result
	} 

```

```{R}
# Function to apply chisq.test() to tibble groups
diffCPS.test = function(tm, cn, rc) {
	cont.t = data.frame(time = tm, cpsNo = cn, readCount = rc) %>%
		spread(time, readCount)
	return(chisq.test(cont.t[, -1])$p.value)}
TSI.cor.test=function(cps,tm,rc){
  cont.t=data.frame(CPS=cps,Time=tm,Readcounts=rc)%>%
    spread(CPS,Readcounts)
mat=as.matrix(cont.t[,-1])
t=matrix(as.vector(mat),byrow=F,ncol=ncol(mat),dimnames = list("time"=rownames(mat),
                                                                "ALEindex"=as.character(c(colnames(mat)))
))

#t=prop.table(t,margin=1)*100
t=as.table(round(t,0)) #when you use mean.rc
x=pears.cor(t,cscore=1:ncol(t),rscore=1:nrow(t))
#lt <- lbl_test(t, scores = list("utrlen"=1:ncol(t),"time"=1:nrow(t)))
  return(x[1])
}




```
#code below used only for tandem UTR switch
```{R}
##edited TSI.cor.test

TSI.cor.test=function(cps,tm,rc,cpsutrlen){
  cont.t=data.frame(CPS=cps,Time=tm,Readcounts=rc)%>%
    spread(CPS,Readcounts)
  
  df=data.frame(CPS=cpsutrlen,Time=tm)%>%
    separate(CPS,c("Tandem","utrlen"),sep=":")%>%
    select(-Time)%>%arrange(Tandem,utrlen)%>%
    distinct()%>%mutate(utrlen=as.numeric(utrlen))%>%
  mutate(Tandem=as.numeric(Tandem))#%>%spread(UTRlen,Readcounts)
vec=df$Tandem
names(vec)=df$utrlen
mat=as.matrix(cont.t[,-1])
t=matrix(as.vector(mat),byrow=F,ncol=ncol(mat),dimnames = list("time"=rownames(mat),
                                                                "CPSindex"=as.character(c(colnames(mat)))
))

t=as.table(round(t,0)) 
x=pears.cor(t,cscore=as.numeric(names(vec)),rscore=1:nrow(t))
  return(x[1])
}


```

1. Load temporal 3p-seq data with annotated ALE, and tandem UTR
```{R}
cps<-read.table("3pseq.all.temporalCPSsamples.ALE.tandemUTR.UTRlen.annotation.txt",header=T,stringsAsFactors = F)

```
2. ALE switch test.
select genes with multiple ALEs
2.1 Distal ALE is a preferred isoform in the cells.
```{R}
ale.sum=cps%>%group_by(hgnc,ALE,time)%>%summarise(s=sum(readcount))%>%ungroup()
ale.summary=ale.sum%>%group_by(hgnc)%>%summarise(alecount=n_distinct(ALE))%>%ungroup()%>%
  filter(alecount>1&alecount<=3)%>%select(hgnc,alecount)%>%inner_join(ale.sum,by="hgnc")%>%ungroup()%>%group_by(time,hgnc)%>%mutate(PSI=s/sum(s))
pdf("Fig3B_ALE expression_uponLPS.pdf",width=5,height=3)
ggplot(ale.summary,aes(x=as.factor(time),fill=as.factor(ALE),y=PSI))+
  geom_boxplot()+facet_wrap(.~alecount)+xlab("time upon LPS")+ylab("PSI")+scale_fill_manual(values=yk.col(5)[c(2,4,5)])+labs(fill="ALE index")+theme_bw()
dev.off()

```



```{R}
library(gridExtra)
yk.col = function(n = 5, sat = 0.7, lum = 1) {
  col = c("royalblue4", "deepskyblue3", "turquoise3",
          "grey62", "goldenrod2", "orange2", "orangered2") %>%
    col2rgb %>%
    rgb2hsv
  return(colorRampPalette(hsv(col[1,], col[2,] * sat,
                              col[3,] * lum))(n)) }

par(mar = c(2,2,1,1), mgp = c(1.5,0.5,0))
col = yk.col(5)[c(1,2,4,5)] 
cdf_common = ggplot() +
  scale_color_manual(values = col) +
  theme_bw() +
  xlim(0,1) +
  ylab("Cumulative fraction")


```

Draw CDF plot
```{R}
all=ale.summary%>%filter((alecount==3&ALE==3)|(alecount==2&ALE==2))

pdf("Fig3C_cdf_PSI_Distal_upon LPS.pdf",width=5,height=3)
cdf_common +
  stat_ecdf(data = all,
            aes(x = PSI,
                col = as.factor(time)),
            geom = "step") +
  labs( title =expression("Genes with multiple ALEs" ),color ="Time") +facet_grid(.~alecount)+
xlab("PSI of distal ALE isoform")
dev.off()
```


Global Switch in ALEs didn't occur upon LPS stimulation
```{R}
all=list()

for (i in 2:3){
ale.summary=ale.sum%>%group_by(hgnc)%>%summarise(alecount=n_distinct(newALE))%>%ungroup()%>%
  filter(alecount==i)%>%select(hgnc)%>%inner_join(ale.sum,by="hgnc")
ale.summary2=ale.summary%>%spread(newALE,s)%>%mutate(Distal=.[[i+2]])%>%
  select(time,Distal,hgnc)%>%
  mutate(ALE_count=i)
all[[i]]=ale.summary2
}
all=all%>%bind_rows()
cat("number of genes with two ALEs",unlist(all%>%group_by(ALE_count)%>%summarise(n_distinct(hgnc))))
ggplot(all,aes(x=as.factor(time),y=Distal,fill=as.factor(ALE_count)))+
  geom_boxplot(notch = T)+
  
  facet_grid(.~ALE_count)


ale.summary%>%group_by(alecount)%>%summarise(n_distinct(hgnc))

```

2.3 chisq.test and PSI



ALE switch
dPSI>0.1
#tandem 3'UTR switch 
```{R}
ale.sum=cps%>%group_by(hgnc,ALE,time)%>%summarise(s=sum(readcount))
final2=ale.sum%>%group_by(hgnc)%>%summarise(alecount=n_distinct(ALE))%>%ungroup()%>%
  filter(alecount>1)%>%select(hgnc)%>%inner_join(ale.sum,by="hgnc")
final2%>%summarise(n_distinct(hgnc)) #2690 genes with multiple ALE
id20=final2%>%group_by(hgnc)%>%summarise(s=sum(s))%>%filter(s>20)%>%select(hgnc) #80 removed
final2=final2%>%inner_join(id20,by="hgnc")
```


```{r}
timelist=list(c(0,1),c(1,2),c(0,2),c(2,4),c(1,4),c(0,4))
result=list()
for ( i in 1:6){
 final.tmp=final2%>%filter(time%in%timelist[[i]]) 

cr.pv = final.tmp %>%
	ungroup() %>%
	group_by(hgnc) %>%
	summarise(pval = diffCPS.test(cn=ALE,tm=time,rc=s))

cr.pv = cr.pv %>%
	mutate(fdr = p.adjust(pval, method = "fdr")) %>%
	mutate(pval = ifelse(is.na(pval), 1, pval)) %>%
	mutate(fdr = ifelse(is.na(fdr), 1, fdr))

cr.pv2 = final.tmp %>%
	ungroup() %>%
	group_by(hgnc) %>%
	summarise(pcor = TSI.cor.test(cps=ALE,tm=time, rc=s))

cr.pv=cr.pv%>%inner_join(cr.pv2,by="hgnc")#%>%inner_join(cr.pv3, by="hgnc")

cr.pv=cr.pv%>%filter_all(all_vars(!is.nan(.)))
result[[i]]<-cr.pv
}
time=c("0-1h","1-2h","0-2h","2-4h","1-4h","0-4h")
for (i in 1:6){
 result[[i]]=result[[i]]%>%mutate(type=rep(time[i]) )
}

final=result%>%reduce(bind_rows)
write.table(final,"ALEswitch.chisqure.test.result.txt",col.names = T,row.names = F,quote=F )
final%>%
  filter(fdr<0.01)%>%
  filter(abs(pcor)>0.1)%>%mutate(cor=ifelse(pcor>0.1,"distal","proximal"))%>%group_by(cor)%>%select(hgnc,cor,type)%>%distinct()%>%summarise(n_distinct(hgnc))
#
final%>%
  filter(fdr<0.01)%>%
  filter(abs(pcor)>0.1)%>%mutate(cor=ifelse(pcor>0.1,"distal","proximal"))%>%group_by(cor)%>%select(hgnc,cor,type)%>%distinct()%>%group_by(hgnc)%>%summarise(n=n_distinct(cor))%>%filter(n==2)#4 detected for switching to both directions.

```


```{r}
final%>%filter(fdr<0.01)%>%filter(abs(pcor)>0.1)%>%summarise(n_distinct(hgnc))
cat("number of hgnc:ALE with switch in tandem3'UTR")
final%>%filter(fdr<0.01)%>%filter(abs(pcor)>0.1)%>%ungroup()%>%summarise(n_distinct(hgnc)) #38
final%>%filter(fdr<0.01)%>%filter(abs(pcor)>0.1)%>%ungroup()%>%mutate(cor=ifelse(pcor>0.1,"distal","proximal"))%>%group_by(cor)%>%summarise(n_distinct(hgnc))
write.table(final%>%
  filter(fdr<0.01)%>%
  filter(abs(pcor)>0.1)%>%mutate(cor=ifelse(pcor>0.1,"distal","proximal"))%>%group_by(cor)%>%select(hgnc,cor,type,pval,pcor)%>%distinct(),"ALEswitch.chisqure.test.result_passed.txt",col.names = T,row.names = F,quote=F )
t=final%>%group_by(type)%>%filter(fdr<0.01)%>%filter(abs(pcor)>0.1)%>%mutate(direction=ifelse(pcor>0.1,"distal","proximal"))%>%group_by(type,direction)%>%summarise(significant=n())
#t[10,]=c("1-2h","distal",0)
#t[9,]=c("2-4h","distal",0)
t=t%>%ungroup()%>%complete(type,nesting(direction),fill =list(significant=0))%>%mutate(significant=as.numeric(significant))

pdf("Fig3F_hgnc_ALEswitch.pdf",width=4,height=3)
ggplot(t,aes(x=type,y=significant,fill=direction,group=direction))+
  geom_bar(stat="identity")+theme_bw()+xlab("Comparison")+
theme_bw()+ylab("Number of ALE shift")+xlab("Comparison")+scale_fill_manual(values=yk.col(5)[c(4,2)])

dev.off()
final%>%filter(fdr<0.05)%>%filter(abs(pcor)>0.1)%>%ungroup()%>%filter(hgnc=="LRCH4")

```

```{R}
write.table(
final%>%filter_all(all_vars(!is.nan(.)))%>%filter(fdr<0.01)%>%filter((pcor)>(0.1))%>%select(hgnc)%>%distinct(),"TSI.result.switchtoDistal.ALE.txt",col.names = F,row.names = F,quote=F,sep="\t")

write.table(
final%>%filter_all(all_vars(!is.nan(.)))%>%filter(fdr<0.01)%>%filter((pcor)<(-0.1))%>%select(hgnc)%>%distinct(),"TSI.result.switchtoProximal.ALE.txt",col.names = F,row.names = F,quote=F,sep="\t")

write.table(cps%>%select(hgnc)%>%arrange(hgnc)%>%distinct(),"ALEswitch.analysis.backgroundhgnc.n.10497.txt",col.names = F,row.names = F,quote=F,sep="\t") 
cps%>%select(hgnc)%>%arrange(hgnc)%>%distinct()%>%nrow
```

Example genes.

TSI index
```{R}
# install.packages("devtools")
#devtools::install_github("haleyjeppson/ggmosaic")
library(ggmosaic)
#install.packages("vcd")
library("vcd")
# plot just a subset of the table
f=final2%>%ungroup()%>%filter(hgnc=="PSMD4")%>%select(newALE,time,s)%>%spread(newALE,s)#%>%
f=final2%>%ungroup()%>%filter(hgnc=="NUB1")%>%select(newALE,time,s)%>%spread(newALE,s)
f=ale.sum%>%ungroup()%>%filter(hgnc=="CLTA")%>%select(ALE,time,s)%>%spread(ALE,s)
f=ale.sum%>%ungroup()%>%filter(hgnc=="LRCH4")%>%select(ALE,time,s)%>%spread(ALE,s)

f=f%>%mutate(time=as.numeric(time))
f
mat=as.matrix(f[c(1:4),-1])
rownames(mat)=c(0,1,2,4)
t=matrix(as.vector(mat),byrow=F,ncol=ncol(mat),dimnames = list("time"=rownames(mat),
                                                                "ALE index"=colnames(mat)))

#t=prop.table(t,margin=1)*100
t=as.table(round(t))
t

#lt <- lbl_test(t, scores = list("Relative Position from the reference poly(A) site"=1:ncol(t),"time"=1:nrow(t)))
cmh_test(t)

spineplot(t)
library("graphics")


pdf("Fig3I_LRCH4_SWITCHtoProximalALE_mosaic.pdf",width=3.5,height=4)
    mosaicplot((t), shade = F, las=1,cex.axis = 0.66,
           main = "Switch in ALEs for LRCH4",color = yk.col(ncol(t))[1:ncol(t)],labeling_args = list(
     gp_labels = gpar(fontsize = 12, fontface = 3),
     gp_varnames = gpar(fontsize = 16, fontface = 2)))

dev.off()


pdf("LRCH4_horizontal.pdf",width=3,height=4)
assocplot(t, col=c("red","blue"),space =0.5)
dev.off()



pdf("CLTA_SWITCHtoDistalALE_mosaic.pdf",width=3.5,height=4)
    mosaicplot((t), shade = F, las=1,cex.axis = 0.66,
           main = "Switch in ALEs for CLTA",color = yk.col(ncol(t))[1:ncol(t)],labeling_args = list(
     gp_labels = gpar(fontsize = 12, fontface = 3),
     gp_varnames = gpar(fontsize = 16, fontface = 2)))

dev.off()


pdf("CLTA_horizontal.pdf",width=3,height=4)
assocplot(t, col=c("red","blue"),space =0.5)
dev.off()

```

Switch in ALE vs RNA level change.
0h vs 1h
```{R}
test=final%>%filter(type=="0-1h")%>%filter(pcor!=1)
test.rna=ale.sum%>%group_by(hgnc,time)%>%summarise(sum.rc=sum(s))%>%spread(time,sum.rc)%>%mutate(dRNA=log2(`1`/`0`))%>%select(hgnc,dRNA)
test.tmp=test%>%inner_join(test.rna,by="hgnc")
test.tmp=test.tmp%>%
  #filter(fdr<0.05&abs(pcor)>0.1)%>%
  mutate(col=ifelse(fdr<0.01&pcor>0.1,"To Distal",ifelse(fdr<0.01&pcor<(-0.1),"To Proximal","No change")))%>%mutate(col=factor(col,levels=c("No change","To Proximal","To Distal")))

#test.tmp=test.tmp%>%filter(pval<0.01)%>%mutate(col=ifelse(pcor>0.1,"To Distal",ifelse(pcor<(-0.1),"To Proximal","No change")))
pdf("dRNA vs ALE switch0.1.pdf",width=5,height=4)
ggplot(test.tmp,aes(x=pcor,y=dRNA,col=col))+geom_jitter (aes(size=-log10(fdr)))+scale_colour_manual(values=c("black","red","blue"))+theme_bw()+xlab("ALE switch index")+ylab("log2(LPS 1h/LPS 0h)")+xlim(c(-1,1))+ylim(c(-6,6))
dev.off()
pdf("Fig3G_dRNA vs ALE switch_boxplot0.1.pdf",width=4,height=3)
ggplot(test.tmp,aes(x=col,y=dRNA,fill=col))+
  geom_boxplot(notch=T)+
  scale_fill_manual(name="Direction",values=yk.col(5)[c(3,2,4)])+theme_bw()+xlab("ALE switch")+ylab("log2(LPS 1h/LPS 0h)")
dev.off()
t.test(test.tmp$dRNA[test.tmp$col=="To Proximal"],test.tmp$dRNA[test.tmp$col=="To Distal"])
```

0h vs 2h
```{R}
test=final%>%filter(type=="0-2h")%>%filter(pcor!=1)
test.rna=ale.sum%>%group_by(hgnc,time)%>%summarise(sum.rc=sum(s))%>%spread(time,sum.rc)%>%mutate(dRNA=log2((`2`+1)/(`0`+1)))%>%select(hgnc,dRNA)
test.tmp=test%>%inner_join(test.rna,by="hgnc")
test.tmp=test.tmp%>%
  #filter(fdr<0.05&abs(pcor)>0.1)%>%
  mutate(col=ifelse(fdr<0.01&pcor>0.1,"To Distal",ifelse(fdr<0.01&pcor<(-0.1),"To Proximal","No change")))%>%mutate(col=factor(col,levels=c("No change","To Proximal","To Distal")))

#test.tmp=test.tmp%>%filter(pval<0.01)%>%mutate(col=ifelse(pcor>0.1,"To Distal",ifelse(pcor<(-0.1),"To Proximal","No change")))
pdf("dRNA vs ALE switch0.2.pdf",width=5,height=4)
ggplot(test.tmp,aes(x=pcor,y=dRNA,col=col))+geom_jitter (aes(size=-log10(fdr)))+scale_colour_manual(values=c("black","red","blue"))+theme_bw()+xlab("ALE switch index")+ylab("log2(LPS 2h/LPS 0h)")+xlim(c(-1,1))+ylim(c(-6,6))
dev.off()
pdf("Fig3G_dRNA vs ALE switch_boxplot0.2.pdf",width=4,height=3)
ggplot(test.tmp,aes(x=col,y=dRNA,fill=col))+
  geom_boxplot(notch=T)+
  scale_fill_manual(name="Direction",values=yk.col(5)[c(3,2,4)])+theme_bw()+xlab("ALE switch")+ylab("log2(LPS 2h/LPS 0h)")
dev.off()
t.test(test.tmp$dRNA[test.tmp$col=="To Proximal"],test.tmp$dRNA[test.tmp$col=="To Distal"])
```

0h vs 4h

```{R}
test=final%>%filter(type=="0-4h")%>%filter(pcor!=1)
test.rna=ale.sum%>%group_by(hgnc,time)%>%summarise(sum.rc=sum(s))%>%spread(time,sum.rc)%>%mutate(dRNA=log2((`4`+1)/(`0`+1)))%>%select(hgnc,dRNA)
test.tmp=test%>%inner_join(test.rna,by="hgnc")
test.tmp=test.tmp%>%
  #filter(fdr<0.05&abs(pcor)>0.1)%>%
  mutate(col=ifelse(fdr<0.01&pcor>0.1,"To Distal",ifelse(fdr<0.01&pcor<(-0.1),"To Proximal","No change")))%>%mutate(col=factor(col,levels=c("No change","To Proximal","To Distal")))

pdf("Fig3G_dRNA vs ALE switch_boxplot0.4.pdf",width=4,height=3)
ggplot(test.tmp,aes(x=col,y=dRNA,fill=col))+
  geom_boxplot(notch=T)+
  scale_fill_manual(name="Direction",values=yk.col(5)[c(3,2,4)])+theme_bw()+xlab("ALE switch")+ylab("log2(LPS 4h/LPS 0h)")
dev.off()
t.test(test.tmp$dRNA[test.tmp$col=="To Proximal"],test.tmp$dRNA[test.tmp$col=="To Distal"])
test.tmp%>%group_by(col)%>%summarise(n())

```


