---
title: "F1B_Spikein CDF"
author: "Yeonui+Kwak"
date: "4/7/2020"
output: html_document
---

1.Set a working directory
```{R}
setwd("~/Desktop/Publication_Mar/Fig1/Tmp13")

```
#set up color
```{r}
library(RColorBrewer)
library(dplyr)
library(tidyr)

yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }
```

```{r}
par(mar = c(2,2,1,1), mgp = c(1.5,0.5,0))
col = yk.col(7)[c(2,3,5,7)] 
```




2. Load the required libraries.
```{R}

library(dplyr)
library(tidyr)
library(ggplot2)
library(gridExtra)
#source("scatterPlot.R")
```


3. Load data

```{R}
dat<-read.table("THP1_TED_0HR_SpikeIn.bedGraph")
dat<-read.table("THP1_TED_1HR_SpikeIn.bedGraph")
dat<-read.table("THP1_TED_2HR_SpikeIn.bedGraph")
dat<-read.table("THP1_TED_4HR_SpikeIn.bedGraph")
dat<-read.table("TED_THP1_rep2_PE_LPS1hr_SpikeIn.bedGraph")
head(dat)

```
4. Data manipulation
```{r}
colnames(dat)=c("Species","Start","End","Freq")
dat=dat%>%select(-Start)%>%mutate(Species=factor(Species, levels=c("A40","A80","A120","A160")))
table(dat$Species)
```

```{R}
pos.mat=NULL
for (species in c("A40","A80","A120","A160")){
dat2=dat%>%filter(Species ==species)
mat=rep(0,690)
for (i in 1:nrow(dat2)){
 mat[dat2[i,2]]=dat2[i,3]
} 
pos.mat=rbind(pos.mat,mat)
}
str(pos.mat)
```

```{R}
row.names(pos.mat)=c("A40","A80","A120","A160")
trans.pos.mat=t(pos.mat)
t<-cbind(pos=1:690,trans.pos.mat)
mat2=t
mat2_long=data.frame(mat2) %>% gather(key=species,value=readcounts,-pos)%>%mutate(species=factor(species,levels = c("A40","A80","A120","A160")))%>%group_by(species) %>%
  mutate(cum_sum = cumsum(readcounts))%>%mutate(cum_sum=cum_sum/max(cum_sum))%>%mutate(pal=pos-690+300)%>%filter(pal>0)

```
#Before Normalization.
```{r}
pdf("cdf_spike-in Rep_LPS 4hr.pdf",width=3,height=2.5)
  ggplot(mat2_long, aes(x=pal, y=cum_sum, col=species)) + geom_step(size = 0.7)+xlab("PAL(nt)") +
  labs( title = "spike-in PAL standards") +labs(color = "PAL")+scale_color_manual(values = col) +
    theme_bw() +
    ylab("Cumulative fraction")
dev.off()


pdf("withnoise_Rep2_4hour.pdf")
ggplot(mat2_long,aes(pos,readcounts,col=species))+geom_bar(stat="identity")+facet_grid(species ~.,scales = "free")+geom_path()+scale_color_manual(values = col) +
    theme_bw()
dev.off()

```



Make the Distribution sharp by settting threshold.
```{R}
mat2_long=data.frame(mat2) %>% gather(key=species,value=readcounts,-pos)%>%mutate(species=factor(species,levels = c("A40","A80","A120","A160")))%>%group_by(species) %>%
  mutate(cum_sum = cumsum(readcounts))%>%mutate(cum_sum=cum_sum/max(cum_sum))%>%mutate(pal=pos-690+300)%>%filter(pal>0)
threshold=c()
#A40
t=mat2_long %>% filter(species=="A40")
i=max(t$readcounts)
while(i>0){
  
if ((sum(t$readcounts[t$readcounts>=i])/sum(t$readcounts))>=0.6){
 threshold=c(threshold,i)
  t$readcounts[t$readcounts<i]=0
  break()
}
  i=i-1
}

mat2_long$readcounts[mat2_long$species=="A40"]=t$readcounts

#A80
t=mat2_long %>% filter(species=="A80")
i=max(t$readcounts)
while(i>0){
  
if ((sum(t$readcounts[t$readcounts>=i])/sum(t$readcounts))>=0.6){
 threshold=c(threshold,i)
  t$readcounts[t$readcounts<i]=0
  break()
}
  i=i-1
}
mat2_long$readcounts[mat2_long$species=="A80"]=t$readcounts
```

```{r}
#A120
t=mat2_long %>% filter(species=="A120")
i=max(t$readcounts)
while(i>0){
  
if ((sum(t$readcounts[t$readcounts>=i])/sum(t$readcounts))>=0.6){
 threshold=c(threshold,i)
  t$readcounts[t$readcounts<i]=0
  break()
}
  i=i-1
}
mat2_long$readcounts[mat2_long$species=="A120"]=t$readcounts
#A160
t=mat2_long %>% filter(species=="A160")
i=max(t$readcounts)
while(i>0){
  
if ((sum(t$readcounts[t$readcounts>=i])/sum(t$readcounts))>=0.6){
 threshold=c(threshold,i)
  t$readcounts[t$readcounts<i]=0
  break()
}
  i=i-1
}

mat2_long$readcounts[mat2_long$species=="A160"]=t$readcounts
threshold
```
1. After making a distribution sharp, Draw a CDF plot.
```{R}
mat2_long=mat2_long%>%group_by(species) %>%
  mutate(cum_sum = cumsum(readcounts))%>%mutate(cum_sum=cum_sum/max(cum_sum))

pdf("F1B_Sup_Rep1_0hr_CDF.pdf",width=3,height=2.5)
  ggplot(mat2_long, aes(x=pal, y=cum_sum, col=species)) + geom_step(size = 0.7)+xlab("PAL (nt)") +
  labs( title = "spike-in PAL standards") +labs(color = "PAL")+scale_color_manual(values = col) +
    theme_bw() +
    ylab("Cumulative fraction")
dev.off()


pdf("F1B_Sup_Rep1_0hr_Hist.pdf.pdf")
ggplot(mat2_long,aes(pos,readcounts,col=species))+geom_bar(stat="identity")+facet_grid(species ~.,scales = "free")+geom_path()+scale_color_manual(values = col) +xlab("PAL (nt)")+ylab("Readcounts")+theme_bw()
dev.off()

write.table(mat2_long,"Rep1_SpikeIn_0hr.final.txt")
```







```{R}
#merge in a grid
require(cowplot)
pdf("A120.pdf")
before.120<-ggplot(t,aes(pos,readcounts))+geom_bar(stat="identity", fill="darkblue")
after.120<-ggplot(t2,aes(pos,readcounts))+geom_bar(stat="identity", fill="red")

pdf("A160.pdf")
before.160<-ggplot(t,aes(pos,readcounts))+geom_bar(stat="identity", fill="darkblue")
after.160<-ggplot(t2,aes(pos,readcounts))+geom_bar(stat="identity", fill="red")
plot_grid(before.160, after.160, labels = c('A', 'B'),ncol = 1, align = 'v')
dev.off()

pdf("before_all.pdf")
plot_grid(before.40,before.80,before.160, labels = c('A', 'B','C'),ncol = 1, align = 'v')
dev.off()
pdf("After_all.pdf")
plot_grid(after.40,after.80,after.160, labels = c('A', 'B','C'),ncol = 1, align = 'v')
dev.off()
pdf("histogram.pdf")
ggplot(mat2_long,aes(readcounts,fill=species))+geom_histogram(position="dodge")+xlab("Number of mapped reads")+ylab("Number of genome positions")
dev.off()



