---
title: "F5S8: ks.test and count the number of genes with PAL changes.:new transcript annotation.(6.12 .2020 ver.)"
author: "Yeonui+Kwak"
date: "6/12/2020"
output: html_document
---

1.Set a working directory
```{R}
setwd("~/Desktop/Publication_Mar/Fig5/Tmp8")

```
2.Load the required packages,and set up colorcode.
```{r}
library(RColorBrewer)
#nstall.packages("raster")
library(raster)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }
```
3. Load the files for ks.test
Spike-in Reference.
#normalize by spike-in
ted$rep1_0=ted$rep1_0+13
ted$rep1_1=ted$rep1_1+15
ted$rep1_2=ted$rep1_2+13
ted$rep1_4=ted$rep1_4+19
ted$rep2_0=ted$rep2_0+16
ted$rep2_1=ted$rep2_1+13
ted$rep2_2=ted$rep2_2+12
ted$rep2_4=ted$rep2_4+8
ted=as_tibble(ted)


3.Statistical Test to identify genes with Significant PAL changes.
3.1 Load the TED-seq position matrix
```{r}
t0<-read.csv("tedseq.custom.rep1.0hr.txt",sep="\t",header=F, stringsAsFactors = F)
t1<-read.csv("tedseq.custom.rep1.1hr.txt",sep="\t",header=F, stringsAsFactors = F)
t2<-read.csv("tedseq.custom.rep1.2hr.txt",sep="\t",header=F, stringsAsFactors = F)
t4<-read.csv("tedseq.custom.rep1.4hr.txt",sep="\t",header=F, stringsAsFactors = F)
```
3.1 Load the TED-seq position matrix
```{r}
start=Sys.time()
t2.0<-read.csv("tedseq.custom.rep2.0hr.txt",sep="\t",header=F, stringsAsFactors = F)
t2.1<-read.csv("tedseq.custom.rep2.1hr.txt",sep="\t",header=F, stringsAsFactors = F)
t2.2<-read.csv("tedseq.custom.rep2.2hr.txt",sep="\t",header=F, stringsAsFactors = F)
t2.4<-read.csv("tedseq.custom.rep2.4hr.txt",sep="\t",header=F, stringsAsFactors = F)
end=Sys.time()
cat("Running time for loading the files was",end-start,"mins") #2mins
```

3.2 Put four matrix to each variable.
Rep1.
```{r}
ncol(First)
First=t0
Second=t1
Third=t2
Four=t4
```
Rep2.
```{r}
First=t2.0
Second=t2.1
Third=t2.2
Four=t2.4
```

3.3 filter transcripts with more than 50 read counts.
```{R}
start=Sys.time()
dat2=as_tibble(First);dat2$sum=rowSums(dat2[203:502])
d0h=dat2%>%filter(sum>50)%>%
  select(-sum)%>%
  select(1,203:502)

dat2=as_tibble(Second);dat2$sum=rowSums(dat2[203:502])
d1h=dat2%>%filter(sum>50)%>%
  select(-sum)%>%select(1,203:502)

dat2=as_tibble(Third);dat2$sum=rowSums(dat2[203:502])
d2h=dat2%>%filter(sum>50)%>%
  select(-sum)%>%select(1,203:502)
dat2=as_tibble(Four);dat2$sum=rowSums(dat2[203:502])
d4h=dat2%>%filter(sum>50)%>%
  select(-sum)%>%select(1,203:502)
```

innerjoin all samples.
```{R}
m=list(d0h,d1h,d2h,d4h)%>%reduce(inner_join,by="V1")
m2=m%>%separate(V1,c("id","refpos","chr","cpspos"),sep=":")%>%unite("newid",c(id,refpos,cpspos),sep=":")
head(m2)
#filter PAS isoforms without local tandem PAS switch as well as without nearly PAS.
head(final.pas)
final.pas<-read.table("custom.PAL.table.afterRemovinglocalPALswitch.withoutnearbyPAS.txt",header=T,stringsAsFactors = F)
final.pas=final.pas%>%separate(newid,c("id","cpspos"),sep=":")%>%unite("newid",c(id,refpos,cpspos),sep=":")
m=m2%>%filter(newid %in%final.pas$newid)%>%select(-chr)
head(m)
```

1. Calculate median PAL, and ks.test pvalue after normalizing spike-in values.
```{R}
#rep1
spike_first=13
spike_second=15
spike_third=13
spike_fourth=19

#rep2
spike_first=16
spike_second=13
spike_third=12
spike_fourth=8
```


rep1
```{R}
start=Sys.time()
ks.pval.1=c();ks.pval.2=c();ks.pval.3=c();ks.pval.4=c();ks.pval.5=c();ks.pval.6=c()
dmean.1=c();dmean.2=c();dmean.3=c();dmean.4=c();dmean.5=c();dmean.6=c()

for (i in 1:nrow(m)){
  vec_1=c()
  for (j in 1:300){
    vec_1=c(vec_1,rep(j+spike_first,m[i,j+1]))
  }
  vec_2=c()
  for (k in 1:300){
    vec_2=c(vec_2,rep(k+spike_second, m[i,k+301]))
  }
  vec_3=c()
  for (k in 1:300){
    vec_3=c(vec_3,rep(k+spike_third, m[i,k+601]))
  }
  vec_4=c()
  for (k in 1:300){
    vec_4=c(vec_4,rep(k+spike_fourth, m[i,k+901]))
  }
   ks.pval.1=c(ks.pval.1,ks.test(vec_1,vec_2)$p.value)
   ks.pval.2=c(ks.pval.2,ks.test(vec_1,vec_3)$p.value)
   ks.pval.3=c(ks.pval.3,ks.test(vec_1,vec_4)$p.value)
   ks.pval.4=c(ks.pval.4,ks.test(vec_2,vec_3)$p.value)
   ks.pval.5=c(ks.pval.5,ks.test(vec_2,vec_4)$p.value)
   ks.pval.6=c(ks.pval.6,ks.test(vec_3,vec_4)$p.value)
   dmean.1=c(dmean.1,mean(vec_2)-mean(vec_1))
   dmean.2=c(dmean.2,mean(vec_3)-mean(vec_1))
   dmean.3=c(dmean.3,mean(vec_4)-mean(vec_1))
   dmean.4=c(dmean.4,mean(vec_3)-mean(vec_2))
   dmean.5=c(dmean.5,mean(vec_4)-mean(vec_2))
   dmean.6=c(dmean.6,mean(vec_4)-mean(vec_3))
}
m$`kspval.0:1`=p.adjust(ks.pval.1,method="BH")
m$`kspval.0:2`=p.adjust(ks.pval.2,method="BH")
m$`kspval.0:4`=p.adjust(ks.pval.3,method="BH")
m$`kspval.1:2`=p.adjust(ks.pval.4,method="BH")
m$`kspval.1:4`=p.adjust(ks.pval.5,method="BH")
m$`kspval.2:4`=p.adjust(ks.pval.6,method="BH")

m$`dpal.0:1`=dmean.1
m$`dpal.0:2`=dmean.2
m$`dpal.0:4`=dmean.3
m$`dpal.1:2`=dmean.4
m$`dpal.1:4`=dmean.5
m$`dpal.2:4`=dmean.6
end=Sys.time()
cat("Running time for loading the files was",end-start,"mins")
head(final.pas)
stat.final2=m%>%select(newid,contains("dpal"),contains("kspval"))%>%inner_join(final.pas,by="newid")
head(stat.final2)
write.table(stat.final,"final6086PASisoforms_stattest_meanPAL_rep1.txt",col.names = T,row.names = F,quote = F,sep="\t")
write.table(stat.final2,"final6086PASisoforms_stattest_meanPAL_rep2.txt",col.names = T,row.names = F,quote = F,sep="\t")
```


```{R}

rep1<-read.table("final6086PASisoforms_stattest_meanPAL_rep1.txt", header=T, stringsAsFactors = F)
rep2<-read.table("final6086PASisoforms_stattest_meanPAL_rep2.txt", header=T, stringsAsFactors = F)
head(rep1)
rep1=rep1[,c(1:15,24:28)]
rep1=rep1%>%gather(key=sample,value=value,-hgnc,-ALE,-relpos,-strand,-newid,-Tandem2,-ALEpos,-cluster)%>%separate(sample,c("test","time1","time2"),sep="\\.")%>%unite("comparison",c(time1,time2),sep=":")%>%mutate(rep=1)%>%unite("test_rep",c(test,rep),sep=":")
#colnames(rep1)=c("newid","dpal_1.0.1","dpal_1.0.2","dpal_1.0.4","dpal_1.1.2","dpal_1.1.4","dpal_1.2.4","kspval_1.0.1","kspval_1.0.2","kspval_1.0.4","kspval_1.1.2","kspval_1.1.4","kspval_1.2.4","hgnc","ALE","relpos","strand","Tandem2","ALEpos","cluster")
rep2=rep2[,c(1:15,24:28)]
rep2=rep2%>%gather(key=sample,value=value,-hgnc,-ALE,-relpos,-strand,-newid,-Tandem2,-ALEpos,-cluster)%>%separate(sample,c("test","time1","time2"),sep="\\.")%>%unite("comparison",c(time1,time2),sep=":")%>%mutate(rep=2)%>%unite("test_rep",c(test,rep),sep=":")
head(rep2)
all=bind_rows(rep1,rep2)%>%spread(test_rep,value)
head(all)
```


5. Filter genes that have PAL change using ks.test.
```{R}
#rep1<-read.table("0hand1hPAL_re1_t.test.txt", header=T, stringsAsFactors = F)
#rep2<-read.table("0hand1hPAL_re2_t.test.txt", header=T, stringsAsFactors = F)
rep1<-read.table("final6086PASisoforms_stattest_meanPAL_rep1.txt", header=T, stringsAsFactors = F)
rep2<-read.table("final6086PASisoforms_stattest_meanPAL_rep2.txt", header=T, stringsAsFactors = F)
head(rep1)
rep1.pal=rep1[,c(1,16:23)]%>%mutate(rep=1)
rep2.pal=rep2[,c(1,16:23)]%>%mutate(rep=2)
pal.all=bind_rows(rep1.pal,rep2.pal)
pal.all.mean=pal.all%>%gather(sample,pal,-newid,-rep)%>%separate(sample,c("replicate","time"),sep="_")%>%select(-replicate)%>%group_by(newid,time)%>%summarise(meanpal=mean(pal))
pal.tmp=pal.all.mean%>%spread(time,meanpal)

write.table(pal.tmp,"custom.meanPAL.acrossalltimepoints.txt",col.names = T,row.names = F,sep="\t",quote=F)

```
1) 0 vs 1
```{R}
tmp=as.data.frame(pal.tmp)%>%  mutate(x=`0`,y=`1`)%>%select(newid,x,y)
str(tmp)
tmp=tmp%>%mutate(density = densCols(log(x)/log(2),log(y)/log(2),
                              nbin=512, bandwidth = 0.1,
                              colramp = colorRampPalette(yk.col(5))))

id_=all%>%
  filter(comparison=="0:1")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%select(newid)
all.id[[1]]=id_
cat("total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR <0.01:",nrow(id_))
all%>%
  filter(comparison=="0:1")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%group_by(change)%>%summarise(n())

t2=tmp%>%filter(newid %in% unlist(id_)) %>%select(x,y)#%>%filter(abs(y-x)>=10)
```


6.Draw a scatter plot with genes with PAL change labelled.
```{R}
pdf("F1D_0and1TEDseqscattorplot.new.pdf", width = 3, height = 3)
ggplot(tmp, aes(y = y, x = x)) + 
    geom_point(aes(col = density), size=0.4) + scale_color_identity() +
    theme_bw() + 
    labs(x ="PAL.0h (nt)", y ="PAL.1h (nt)")+xlim(c(0,250))+ylim(c(0,250))+geom_abline(slope = 1, intercept = 10,linetype="dashed")+geom_abline(slope = 1, intercept = -10,linetype="dashed")+geom_abline(slope = 1, intercept = 0,col="grey",size=1, linetype="dotted")+ggtitle("TED-seq : 0h vs 1h")+#ggplot(t2, aes(y = y, x = x))+
    geom_jitter(data = t2,aes(col ="indianred"), size=1)
dev.off()
```

2) 0 vs 2
```{R}
tmp=as.data.frame(pal.tmp)%>%  mutate(x=`0`,y=`2`)%>%select(newid,x,y)
str(tmp)
tmp=tmp%>%mutate(density = densCols(log(x)/log(2),log(y)/log(2),
                              nbin=512, bandwidth = 0.1,
                              colramp = colorRampPalette(yk.col(5))))


id_=all%>%
  filter(comparison=="0:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%select(newid)

cat("total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR <0.01:",nrow(id_))
all%>%
  filter(comparison=="0:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%group_by(change)%>%summarise(n())

t2=tmp%>%filter(newid %in% unlist(id_)) %>%select(x,y)#%>%filter(abs(y-x)>=10)
```
```{R}
pdf("F1D_0and2TEDseqscattorplot.new.pdf", width = 3, height = 3)
ggplot(tmp, aes(y = y, x = x)) + 
    geom_point(aes(col = density), size=0.4) + scale_color_identity() +
    theme_bw() + 
    labs(x ="PAL.0h (nt)", y ="PAL.2h (nt)")+xlim(c(0,250))+ylim(c(0,250))+geom_abline(slope = 1, intercept = 10,linetype="dashed")+geom_abline(slope = 1, intercept = -10,linetype="dashed")+geom_abline(slope = 1, intercept = 0,col="grey",size=1, linetype="dotted")+ggtitle("TED-seq : 0h vs 2h")+#ggplot(t2, aes(y = y, x = x))+
    geom_jitter(data = t2,aes(col ="indianred"), size=1)
dev.off()
```

3) 0 vs 4
```{R}
tmp=as.data.frame(pal.tmp)%>%  mutate(x=`0`,y=`4`)%>%select(newid,x,y)
str(tmp)
tmp=tmp%>%mutate(density = densCols(log(x)/log(2),log(y)/log(2),
                              nbin=512, bandwidth = 0.1,
                              colramp = colorRampPalette(yk.col(5))))


id_=all%>%
  filter(comparison=="0:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%select(newid)
cat("total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR <0.01:",nrow(id_))

all%>%
  filter(comparison=="0:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%group_by(change)%>%summarise(n())

t2=tmp%>%filter(newid %in% unlist(id_)) %>%select(x,y)#%>%filter(abs(y-x)>=10)
```
```{R}
pdf("F1D_0and4TEDseqscattorplot.new.pdf", width = 3, height = 3)
ggplot(tmp, aes(y = y, x = x)) + 
    geom_point(aes(col = density), size=0.4) + scale_color_identity() +
    theme_bw() + 
      labs(x ="PAL.0h (nt)", y ="PAL.4h (nt)")+xlim(c(0,250))+ylim(c(0,250))+geom_abline(slope = 1, intercept = 10,linetype="dashed")+geom_abline(slope = 1, intercept = -10,linetype="dashed")+geom_abline(slope = 1, intercept = 0,col="grey",size=1, linetype="dotted")+ggtitle("TED-seq : 0h vs 4h")+#ggplot(t2, aes(y = y, x = x))+
    geom_jitter(data = t2,aes(col ="indianred"), size=1)
dev.off()
```

4) 1h vs 2h
```{R}
tmp=as.data.frame(pal.tmp)%>%  mutate(x=`1`,y=`2`)%>%select(newid,x,y)

tmp=tmp%>%mutate(density = densCols(log(x)/log(2),log(y)/log(2),
                              nbin=512, bandwidth = 0.1,
                              colramp = colorRampPalette(yk.col(5))))


id_=all%>%
  filter(comparison=="1:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%select(newid)

cat("total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR <0.01:",nrow(id_))
all%>%
  filter(comparison=="1:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%group_by(change)%>%summarise(n())

t2=tmp%>%filter(newid %in% unlist(id_)) %>%select(x,y)#%>%filter(abs(y-x)>=10)
```
```{R}
pdf("F1D_1and2TEDseqscattorplot.new.pdf", width = 3, height = 3)
ggplot(tmp, aes(y = y, x = x)) + 
    geom_point(aes(col = density), size=0.4) + scale_color_identity() +
    theme_bw() + 
      labs(x ="PAL.1h (nt)", y ="PAL.2h (nt)")+xlim(c(0,300))+ylim(c(0,300))+geom_abline(slope = 1, intercept = 10,linetype="dashed")+geom_abline(slope = 1, intercept = -10,linetype="dashed")+geom_abline(slope = 1, intercept = 0,col="grey",size=1, linetype="dotted")+ggtitle("TED-seq : 1h vs 2h")+#ggplot(t2, aes(y = y, x = x))+
    geom_jitter(data = t2,aes(col ="indianred"), size=1)
dev.off()
```

5) 1 h vs 4 h
```{R}
tmp=as.data.frame(pal.tmp)%>%  mutate(x=`1`,y=`4`)%>%select(newid,x,y)
str(tmp)
tmp=tmp%>%mutate(density = densCols(log(x)/log(2),log(y)/log(2),
                              nbin=512, bandwidth = 0.1,
                              colramp = colorRampPalette(yk.col(5))))


id_=all%>%
  filter(comparison=="1:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%select(newid)
cat("total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR <0.01:",nrow(id_))

all%>%
  filter(comparison=="1:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%group_by(change)%>%summarise(n())

t2=tmp%>%filter(newid %in% unlist(id_)) %>%select(x,y)#%>%filter(abs(y-x)>=10)
```
```{R}
pdf("F1D_1and4TEDseqscattorplot.new.pdf", width = 3, height = 3)
ggplot(tmp, aes(y = y, x = x)) + 
    geom_point(aes(col = density), size=0.4) + scale_color_identity() +
    theme_bw() + 
      labs(x ="PAL.1h (nt)", y ="PAL.4h (nt)")+xlim(c(0,250))+ylim(c(0,250))+geom_abline(slope = 1, intercept = 10,linetype="dashed")+geom_abline(slope = 1, intercept = -10,linetype="dashed")+geom_abline(slope = 1, intercept = 0,col="grey",size=1, linetype="dotted")+ggtitle("TED-seq : 1h vs 4h")+#ggplot(t2, aes(y = y, x = x))+
    geom_jitter(data = t2,aes(col ="indianred"), size=1)
dev.off()
```
6) 2 vs 4h
```{R}
tmp=as.data.frame(pal.tmp)%>%  mutate(x=`2`,y=`4`)%>%select(newid,x,y)
tmp=tmp%>%mutate(density = densCols(log(x)/log(2),log(y)/log(2),
                              nbin=512, bandwidth = 0.1,
                              colramp = colorRampPalette(yk.col(5))))


id_=all%>%
  filter(comparison=="2:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%select(newid)
cat("total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR <0.01:",nrow(id_))

all%>%
  filter(comparison=="2:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%group_by(change)%>%summarise(n())

t2=tmp%>%filter(newid %in% unlist(id_)) %>%select(x,y)#%>%filter(abs(y-x)>=10)
```
```{R}
pdf("F1D_2and4TEDseqscattorplot.new.pdf", width = 3, height = 3)
ggplot(tmp, aes(y = y, x = x)) + 
    geom_point(aes(col = density), size=0.4) + scale_color_identity() +
    theme_bw() + 
      labs(x ="PAL.2h (nt)", y ="PAL.4h (nt)")+xlim(c(0,250))+ylim(c(0,250))+geom_abline(slope = 1, intercept = 10,linetype="dashed")+geom_abline(slope = 1, intercept = -10,linetype="dashed")+geom_abline(slope = 1, intercept = 0,col="grey",size=1, linetype="dotted")+ggtitle("TED-seq : 2h vs 4h")+#ggplot(t2, aes(y = y, x = x))+
    geom_jitter(data = t2,aes(col ="indianred"), size=1)
dev.off()
```

total number of transcript isoforms with PAL change gt 10 nt and ks.test FDR<0.01 for all samples.

```{r}
cat("total trascript isoforms tested:",all%>%summarise(n_distinct(newid))%>%unlist)
id_1=all%>%
  filter(comparison=="0:1")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,hgnc,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")

id_2=all%>%
  filter(comparison=="0:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,hgnc,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")
id_3=all%>%
  filter(comparison=="0:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,hgnc,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")
id_4=all%>%
  filter(comparison=="1:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,hgnc,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")
id_5=all%>%
  filter(comparison=="1:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,hgnc,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")
id_6=all%>%
  filter(comparison=="2:4")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.01&`kspval:2`<0.01)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(newid,hgnc,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>=10,"UP",ifelse(mean.dpal<=(-10), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")

all.id=list(id_1,id_2,id_3,id_4,id_5,id_6)%>%reduce(bind_rows)
(all.id)%>%
  #group_by(change)%>%
  summarise(n_distinct(newid))
```

Save PAL dynamics of transcript isoforms with PAL changes (n=1893)
```{R}

write.table(pal.tmp%>%inner_join(all.id%>%arrange(newid)%>%select(newid,hgnc)%>%distinct(),by="newid"),"custom.paldynamics.ofpalchanged isoforms(1893).txt",col.names = T,row.names = F,sep="\t",quote=F)


```

gene with PAL increase between 0 and 1h
FDR<0.05
dPAL>6
```{r}
gene01=all%>%
  filter(comparison=="0:1")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.05|`kspval:2`<0.05)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(hgnc,newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>6,"UP",ifelse(mean.dpal<=(-6), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%filter(mean.dpal>6)%>%separate(newid,c("enst","next"),sep="\\.")

write.table(gene01,"Genelist_PALincrease.0to1h.txt",col.names = T,row.names = F,quote =F,sep="\t")

gene02=all%>%
  filter(comparison=="0:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.05|`kspval:2`<0.05)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(hgnc,newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>6,"UP",ifelse(mean.dpal<=(-6), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%filter(mean.dpal>6)%>%separate(newid,c("enst","next"),sep="\\.")
write.table(gene02,"Genelist_PALincrease.0to2h.txt",col.names = T,row.names = F,quote =F,sep="\t")

gene12=all%>%
  filter(comparison=="1:2")%>%
  filter((`dpal:1`>6&`dpal:2`>6)|(`dpal:1`<(-6)&`dpal:2`<(-6)))%>%
  filter(`kspval:1`<0.05|`kspval:2`<0.05)%>%
  group_by(newid)%>%mutate(mean.dpal=mean(`dpal:1`,`dpal:2`))%>%
  ungroup()%>%
  select(hgnc,newid,mean.dpal)%>%
  mutate(change=ifelse(mean.dpal>6,"UP",ifelse(mean.dpal<=(-6), "DOWN","NOCHANGE")))%>%
  filter(change!="NOCHANGE")%>%filter(mean.dpal>6)%>%separate(newid,c("enst","next"),sep="\\.")
write.table(gene12,"Genelist_PALincrease.1to2h.txt",col.names = T,row.names = F,quote =F,sep="\t")
```
