---
title: 'Step3: Custom poly(A) tail length and 3’UTR length and indexing'
author: "Yeonui+Kwak"
date: "8/8/2020"
output: html_document
---


1.Set a working directory
```{R}
setwd("~/Desktop/Publication_Mar/Final_Fig2/Rcode/")

```

#set up color
```{r}
library(RColorBrewer)
#nstall.packages("raster")
library(raster)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }
```

1.load custom PAL table.

```{R}
#update1 custom rep1 and rep2 mean value was used for analysis!
rep1<-read.table("custom.0h.rep1.PAL.table.txt",header=T,stringsAsFactors = F)
rep2<-read.table("custom.0h.rep2.PAL.table.txt",header=T,stringsAsFactors = F)
#custom transcriptome indexed file.
index<-read.table("transcript.custom.ALE.tandem.index.annotation.txt",header=T,stringsAsFactors = F)
utrlen=read.table("custom.UTRlen.table.txt",header=T,stringsAsFactors = F)
```
```{R}
rep1=rep1%>%unite("newid",c(id,chr,cpspos),sep=":",remove=F)
rep2=rep2%>%unite("newid",c(id,chr,cpspos),sep=":",remove=F)
pal=inner_join(rep1,rep2%>%select(newid,meanpal),by="newid")%>%mutate(meanpal = rowMeans(select(.,meanpal.x,meanpal.y)))
cat("number of transcript isoforms detected in TED",pal%>%summarise(n_distinct(hgnc))%>%unlist())
cat("number of hgnc detected in TED",pal%>%summarise(n_distinct(newid))%>%unlist())
#average of two biological replicates of 0h.
pal.final=pal%>%select(newid,meanpal,relpos)
#indexing the PAL table with ALE and tandem index.
index2=index%>%separate(V4,c("id","refpos","chr","CPSpos"),sep=":",remove=F)%>%unite("newid",c(id,chr,CPSpos),sep=":",remove=F)
utrlen=utrlen%>%separate(V4,c("id","refpos","chr","CPSpos"),sep=":",remove=F)%>%unite("newid",c(id,chr,CPSpos),sep=":",remove=F)
custom=pal.final%>%inner_join(index2,by="newid")%>%select(hgnc,newid,ALE,Tandem,meanpal,CPSpos,V6,relpos)
nrow(custom)
#add custom 3UTR length info
custom=custom%>%inner_join(utrlen%>%select(newid,utrlen),by="newid")
nrow(custom)
```
Output from the code above: nrow:11666 isoforms
```{R}
write.table(custom,"custom.0h.PAL.ALE.Tandem.table.txt",col.names = T,row.names = F,quote=F,sep="\t")
sink("outfile_After custom.0h.PAL.ALE.Tandem.example.txt")

custom%>%filter(hgnc=="UBE2J2")
custom%>%filter(hgnc=="SPSB1")
custom%>%filter(hgnc=="ANTXR1")
custom%>%filter(hgnc=="CD83")
custom%>%filter(hgnc=="SRF")

sink()

```
2. RNA abundance info is added.

```{R}
custom<-read.table("custom.0h.PAL.ALE.Tandem.table.txt",header=T,stringsAsFactors = F)
custom2=custom%>%separate(newid,c("id","chr","CPSpos2"),sep=":")%>%unite("newid",c(id,CPSpos),sep=":",remove = F)%>%select(-CPSpos2)%>%unite("tandemclusterid",c(hgnc,ALE,Tandem),sep=":",remove=F)
head(custom2)
index=custom2%>%select(newid,tandemclusterid)#%>%unite("tandemclusterid",c(hgnc,ALE,Tandem),sep=":")
head(index)
```


```{r}
ratio=read.table("cpspeak.readcount.table.2020.08.07.3UTR1KB.ext.txt",header=T,stringsAsFactors = F)
head(ratio2)
str(ratio)
ratio2=ratio%>%filter(time==0)%>%unite("newid",c(id,CPSpos),sep=":",remove = F)%>%inner_join(index,by="newid")%>%select(newid,tandemclusterid,cps.ratio.perhgnc,readcount)%>%arrange(tandemclusterid,readcount)%>%distinct_at(vars(newid),.keep_all = T)
#Join RNA abundance table to PAL table.
tmp2=custom2%>%ungroup()%>%inner_join(ratio2%>%select(newid,cps.ratio.perhgnc,readcount),by="newid")
head(tmp2)
tmp2%>%filter(hgnc=="CD83")
tmp2=tmp2%>%rename(strand=V6)%>%select(hgnc,newid,id,tandemclusterid,chr,ALE,Tandem,meanpal,CPSpos,strand,relpos,utrlen,cps.ratio.perhgnc,readcount)

#Re index ALE and Tandem
ale.index.tmp=tmp2%>%group_by(hgnc)%>%arrange(ALE)%>%distinct_at(vars(ALE),.keep_all =T)%>%mutate(newALE=rank(ALE))%>%ungroup()%>%unite("hgncale",c(hgnc,ALE),sep=":",remove=F)%>%
  select(hgncale,newALE)%>%inner_join(tmp2%>%unite("hgncale",c(hgnc,ALE),sep=":",remove=F),by="hgncale")%>%select(-hgncale,-ALE)%>%rename(ALE=newALE)

final=ale.index.tmp%>%group_by(hgnc,ALE)%>%arrange(Tandem)%>%distinct_at(vars(Tandem),.keep_all =T)%>%mutate(newTandem=rank(Tandem))%>%ungroup()%>%
  select(tandemclusterid,newTandem)%>%inner_join(ale.index.tmp,by="tandemclusterid")%>%select(-Tandem)%>%rename(Tandem=newTandem)%>%select(-tandemclusterid)%>%unite("tandemclusterid",c(hgnc,ALE,Tandem),sep=":",remove=F)

write.table(final,"final_Fig2TABLEforUTRisoformanalysis.txt",col.names = T,row.names = F,quote=F,sep="\t")


tmp2%>%filter(hgnc=="CD83")
ale.index.tmp%>%filter(hgnc=="CD83")
final%>%filter(hgnc=="CD83")
final%>%filter(hgnc=="CD83")
colnames(final)
nrow(final)
```
Columns
 [1] "tandemclusterid"   "Tandem"            "ALE"               "hgnc"             
 [5] "newid"             "id"                "chr"               "meanpal"          
 [9] "CPSpos"            "strand"            "relpos"            "utrlen"           
[13] "cps.ratio.perhgnc" "readcount"  

2. Local tandem cluster, filter out minor peaks.
Define distance.test
Purpose: When tandem 3'UTR is within 300 nt windows from each other, we will exclude isoforms with minor frequency.
Because it can cause bias that distal tandem UTR always have longer poly(A) because they calculate poly(A) tail length based on the same TED-seq distribution.

```{R}
distance.test=function(cps=CPSpos,ratio=cps.ratio.perhgnc,id=Tandem){
  table=data.frame("cps"=cps,"ratio"=ratio,"id"=id)
  index=c()
  prev=1
  nex=prev+1
  while(prev<nrow(table)&nex<=nrow(table)){
    if(abs(table[prev,1]-table[nex,1])<300){
        if(table[prev,2]>table[nex,2]){
         nex=nex+1 
        }
        else{
          prev=nex
          nex=nex+1
        }
      #index=c(index,prev)
      }
    else{
      index=c(index,table[prev,3])
    prev=nex
    }
  }
  index=c(index,table[prev,3])
  return(str_replace_all(paste(index,",",collapse =""), fixed(" "), ""))
}
```
3. Use the above function, and leave PAL info for non-overlapping PAS isoforms without nearby PAS.
```{R}
#result=#
newCPSid=final%>%group_by(hgnc,ALE)%>%summarise(t=distance.test(cps=CPSpos,ratio=cps.ratio.perhgnc,id=Tandem))%>%
  separate_rows(t,sep=",")%>%filter(t!="")%>%unite("hgnc:ALE:tandem",c(hgnc,ALE,t),sep=":")
cat("without multiple cps within 300 nucleotides.",nrow(newCPSid))
final=final%>%inner_join(newCPSid%>%rename(tandemclusterid=`hgnc:ALE:tandem`), by="tandemclusterid")
nrow(final)
#reindexing again after removing minor peaks.
final2=final%>%group_by(hgnc,ALE)%>%arrange(Tandem)%>%distinct_at(vars(Tandem),.keep_all =T)%>%mutate(newTandem=rank(Tandem))%>%ungroup()%>%
  select(tandemclusterid,newTandem)%>%inner_join(final,by="tandemclusterid")%>%select(-Tandem)%>%rename(Tandem=newTandem)%>%select(-tandemclusterid)%>%unite("tandemclusterid",c(hgnc,ALE,Tandem),sep=":",remove=F)

write.table(final2,"custom.pal.withoutnearbyPAS.txt",col.names = T,row.names = F,sep="\t",quote=F)

final2%>%filter(hgnc=="CD83")

```

