---
title: "F6S5"
author: "Yeonui+Kwak"
date: "6/30/2020"
output: html_document
---

`
1.Set a working directory
```{R}
setwd("~/Desktop/Publication_Mar/Fig6/Tmp5")

```
#set up color
```{r}
library(RColorBrewer)
#nstall.packages("raster")
library(raster)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }
```

3. Load sample
3.1 PAL-RNA
```{R}
final<-read.table("PRO_TED_CPS.readcounttable.6.29.2020.txt",sep="\t",header=T,stringsAsFactors = F)
tnf=final%>%filter(str_detect(id, "ENST0000044926"))#TNFA
final=final%>%filter_if(is.numeric, all_vars(. > 0))
```


5. Select Genes with transcription upregulation.
```{R}
dpro.breaks <- seq(-5,5,1)
tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0>1&PRO_1-PRO_0>=0&PRO_2-PRO_0>=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("PRO"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)

pdf("TxnUP.0.4.dTxn distribution.pdf",width=3,height=2.5)
ggplot(sample,aes(x=dTED_group,y=dPRO,col=dPRO_bin))+geom_quasirandom(size=0.5)+scale_colour_manual(values=yk.col(6)[2:4])+ylab(expression(italic(Delta)*"TXN (PRO-seq log2 FC)"))+labs(title=expression("Txn UP: 0h"%->%"4h"))+theme_bw()+xlab(expression(italic(Delta)*"PAL"))+labs(col=expression(italic(Delta)*"TXN bins"))
dev.off()
```


```{R}
sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%group_by(dTED_group,dCPS_diff)%>%summarise(n())



sample2=sample%>%
  #select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%filter(dCPS_diff=="dCPS_1and2")
col = yk.col(9)[c(2,5)] 
pdf("F5B.Txnup.1.2h.pdf",width=4.5,height=2.5)
ggplot(sample2,aes(x=dTED_group,y=dCPS_value))+
  geom_quasirandom(aes(col=dTED_group),size=1)+
  scale_colour_manual(values=col )+theme_bw()+ylab(expression(italic(Delta)*"RNA(log2 FC (2h/1h))"))+labs(title=expression("Txn UP: 0h"%->%"4h"))+
  labs(col=expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  geom_boxplot(alpha=0.3, color="black", width=.2)+xlab(expression(italic(Delta)*"PAL groups"))+ylim(c(-2,2))
#+geom_beeswarm(size=2,aes(col=dTED_group))
dev.off()

sample2=sample%>%
  #select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%filter(dCPS_diff=="dCPS_2and4")
#col = yk.col(9)[c(5,8)] 
pdf("F5B.Txnup.2.4h.pdf",width=4.5,height=2.5)
ggplot(sample2,aes(x=dTED_group,y=dCPS_value))+
  geom_quasirandom(aes(col=dTED_group),size=1)+
  scale_colour_manual(values=col )+theme_bw()+ylab(expression(italic(Delta)*"RNA(log2 FC (4h/2h))"))+labs(title=expression("Txn UP: 0h"%->%"4h"))+
  labs(col=expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  geom_boxplot(alpha=0.3, color="black", width=.2)+xlab(expression(italic(Delta)*"PAL groups"))+ylim(c(-2.5,2))
#+geom_beeswarm(size=2,aes(col=dTED_group))
dev.off()

col = yk.col(9)[c(2,5)] 
cdf_common = ggplot() +
  scale_color_manual(values = col) +
  theme_bw() +
  xlim(-2,2) +
  ylab("Cumulative fraction")


pdf("F5B.txnup.CDFplot.pdf",width=4.5,height=2.5)
sample2=sample%>%
  #select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))
CPS.labs <- c("RNA : 1h -> 2h", "RNA : 2h -> 4h"); names(CPS.labs) <- c("dCPS_1and2", "dCPS_2and4")
cdf_common +
  stat_ecdf(data = sample2,
            aes(x = dCPS_value,
                col = dTED_group),
            geom = "step") +
  labs( title =expression("Txn UP: 0h"%->%"4h")) +facet_grid(.~dCPS_diff,labeller=labeller( dCPS_diff=CPS.labs))+labs( col =expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  xlab(expression(italic(Delta)*"RNA"))+labs(col=expression(italic(Delta)*"PAL"))#+
dev.off()

pal.labs <- c("PAL : DN", "PAL : NC"); names(pal.labs) <- c("DN", "NC")
col = yk.col(9)[c(2,8)] 
cdf_common = ggplot() +
  scale_color_manual(values = col) +
  theme_bw() +
  xlim(-2,2) +
  ylab("Cumulative fraction")
pdf("F5B.txnup.CDFplot.ver2.pdf",width=4.5,height=2.5)
cdf_common +
  stat_ecdf(data = sample2,
            aes(x = dCPS_value,
                col = dCPS_diff),
            geom = "step") +
  labs( title =expression("Txn UP: 0h"%->%"4h")) +facet_grid(.~dTED_group,labeller=labeller( dTED_group=pal.labs))+labs( col =expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  xlab(expression(italic(Delta)*"RNA"))+labs(col=expression(italic(Delta)*"RNA"))+ scale_color_manual( 
                        labels = c(expression("1h" %->%"2h"), expression("2h" %->% "4h")),values=col)
dev.off()
```


Sampling for FigS4

Transcription upregulation genes.
```{R}
sample
#Validaion no difference in PRO-seq profiles. among TED groups.
sample_=sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
  mutate(.PRO_0=0)%>%mutate(.PRO_1=PRO_1-PRO_0)%>%
  mutate(.PRO_2=PRO_2-PRO_0)%>%mutate(.PRO_4=PRO_4-PRO_0)%>%
  gather("time","txn",starts_with(".PRO_"))%>%
  separate(time,c("PRO","hour"),sep="_")%>%mutate(hour=as.numeric(hour))%>%
  select(dTED_group,hour,txn,newid)
df_summary=sample_%>%group_by(dTED_group,hour)%>%summarise(n=n(),mean=mean(txn),sd=sd(txn),se = sd(txn, na.rm=T)/sqrt(sum(!is.na(txn))))
df_summary$CI_lower <- df_summary$mean + qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
df_summary$CI_upper <- df_summary$mean - qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
#df_summary$CI_lower <- df_summary$mean + df_summary$sd
#df_summary$CI_upper <- df_summary$mean - df_summary$sd

#col = yk.col(10)[c(3,5,7)] 
pdf("F5A.pdf",width=4,height=2.3)
pal.labs <- c("PAL : DN", "PAL : NC", "PAL : UP"); names(pal.labs) <- c("DN", "NC","UP")
ggplot(df_summary, aes(x=hour, y=mean)) +
   geom_line(data=df_summary, aes(x=hour, y=mean,group=dTED_group), size=1, alpha=0.8)+geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper,group=dTED_group,fill=dTED_group), alpha=0.3)+scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+theme_bw()+facet_grid(.~dTED_group,labeller = labeller(dTED_group=pal.labs))+
  #scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+
  ylab("mean(Transcription) ± 95% CI")
dev.off()
```
#  Time-course CPS-seq profiles. among TED groups.

```{r}

tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0>1&PRO_1-PRO_0>=0&PRO_2-PRO_0>=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("CPS_"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)


sample_=sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
mutate(.CPS_0=0)%>%mutate(.CPS_1=log2((CPS_1+1)/(CPS_0+1)))%>%mutate(.CPS_2=log2((CPS_2+1)/(CPS_0+1)))%>%mutate(.CPS_4=log2((CPS_4+1)/(CPS_0+1)))%>%
  gather("time","RNA",starts_with(".CPS_"))%>%separate(time,c("cps","hour"),sep="_")%>%mutate(hour=as.numeric(hour))%>%select(dTED_group,hour,RNA,newid)
df_summary=sample_%>%group_by(dTED_group,hour)%>%summarise(n=n(),mean=mean(RNA),sd=sd(RNA),se = sd(RNA, na.rm=T)/sqrt(sum(!is.na(RNA))))
df_summary$CI_lower <- df_summary$mean + qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
df_summary$CI_upper <- df_summary$mean - qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
#df_summary$CI_lower <- df_summary$mean + df_summary$sd
#df_summary$CI_upper <- df_summary$mean - df_summary$sd


pdf("F5C.pdf",width=4,height=2.3)
pal.labs <- c("PAL : DN", "PAL : NC", "PAL : UP"); names(pal.labs) <- c("DN", "NC","UP")

ggplot(df_summary, aes(x=hour, y=mean)) +
   geom_line(data=df_summary, aes(x=hour, y=mean,group=dTED_group), size=1, alpha=0.8)+geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper,group=dTED_group,fill=dTED_group), alpha=0.3)+theme_bw()+
  facet_grid(.~dTED_group, labeller = labeller(dTED_group=pal.labs))+
scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+ylab("mean(RNA) ± 95% CI")

dev.off()

```

PAL time-course
```{R}


sample=tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0>1&PRO_1-PRO_0>=0&PRO_2-PRO_0>=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("TED_"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)


sample_=sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO,-dTED_value,-dCPS_1and2,-dCPS_2and4)%>%
mutate(.TED_0=0)%>%
  mutate(.TED_1=TED_1-TED_0)%>%
  mutate(.TED_2=TED_2-TED_0)%>%
  mutate(.TED_4=TED_4-TED_0)%>%
  gather("time","PAL",starts_with(".TED_"))%>%separate(time,c("ted","hour"),sep="_")%>%mutate(hour=as.numeric(hour))%>%select(dTED_group,hour,PAL,newid)
df_summary=sample_%>%group_by(dTED_group,hour)%>%summarise(n=n(),mean=mean(PAL),sd=sd(PAL),se = sd(PAL, na.rm=T)/sqrt(sum(!is.na(PAL))))
df_summary$CI_lower <- df_summary$mean + qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
df_summary$CI_upper <- df_summary$mean - qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
#df_summary$CI_lower <- df_summary$mean + df_summary$sd
#df_summary$CI_upper <- df_summary$mean - df_summary$sd


pdf("F5C.PALtimecourse.pdf",width=4,height=2.3)
pal.labs <- c("PAL : DN", "PAL : NC", "PAL : UP"); names(pal.labs) <- c("DN", "NC","UP")

ggplot(df_summary, aes(x=hour, y=mean)) +
   geom_line(data=df_summary, aes(x=hour, y=mean,group=dTED_group), size=1, alpha=0.8)+geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper,group=dTED_group,fill=dTED_group), alpha=0.3)+theme_bw()+facet_grid(.~dTED_group, labeller = labeller(dTED_group=pal.labs))+scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+ylab("mean(PAL) ± 95% CI")

dev.off()



```


#5. Select Genes with transcription Downregulation
```{R}
dpro.breaks <- seq(-5,5,0.5)
tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0<(-1)&PRO_1-PRO_0<=0&PRO_2-PRO_0<=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("PRO"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)

pdf("TxnUP.0.4.DN.dTxn distribution.pdf",width=3,height=2.5)
ggplot(sample,aes(x=dTED_group,y=dPRO,col=dPRO_bin))+geom_quasirandom(size=0.5)+scale_colour_manual(values=yk.col(5)[1:4])+ylab(expression(italic(Delta)*"TXN (PRO-seq log2 FC)"))+labs(title=expression("Txn DN: 0h"%->%"4h"))+theme_bw()+xlab(expression(italic(Delta)*"PAL"))+labs(col=expression(italic(Delta)*"TXN bins"))
dev.off()
```


```{R}
sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%group_by(dTED_group,dCPS_diff)%>%summarise(n())



sample2=sample%>%
  #select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%filter(dCPS_diff=="dCPS_1and2")
col = yk.col(9)[c(2,5)] 
pdf("F5B.txndn.1.2h.pdf",width=4.5,height=2.5)
ggplot(sample2,aes(x=dTED_group,y=dCPS_value))+
  geom_quasirandom(aes(col=dTED_group),size=1)+
  scale_colour_manual(values=col )+theme_bw()+ylab(expression(italic(Delta)*"RNA(log2 FC (2h/1h))"))+labs(title=expression("Txn DN: 0h"%->%"4h"))+
  labs(col=expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  geom_boxplot(alpha=0.3, color="black", width=.2)+xlab(expression(italic(Delta)*"PAL groups"))+ylim(c(-2,1))
#+geom_beeswarm(size=2,aes(col=dTED_group))
dev.off()

sample2=sample%>%
  #select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%filter(dCPS_diff=="dCPS_2and4")
col = yk.col(9)[c(2,5,8)] 
pdf("F5B.txndn.2.4h.pdf",width=4.5,height=2.5)
ggplot(sample2,aes(x=dTED_group,y=dCPS_value))+
  geom_quasirandom(aes(col=dTED_group),size=1)+
  scale_colour_manual(values=col )+theme_bw()+ylab(expression(italic(Delta)*"RNA(log2 FC (4h/2h))"))+labs(title=expression("Txn DN: 0h"%->%"4h"))+
  labs(col=expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  geom_boxplot(alpha=0.3, color="black", width=.2)+xlab(expression(italic(Delta)*"PAL groups"))+ylim(c(-2,1))
#+geom_beeswarm(size=2,aes(col=dTED_group))
dev.off()
```


```{R}
col = yk.col(9)[c(2,5)] 
cdf_common = ggplot() +
  scale_color_manual(values = col) +
  theme_bw() +
  xlim(-2,2) +
  ylab("Cumulative fraction")


pdf("F5B.txndn.CDFplot.pdf",width=4.5,height=2.5)
sample2=sample%>%
  #select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
 gather("dCPS_diff","dCPS_value",starts_with("dCPS"))
CPS.labs <- c("RNA : 1h -> 2h", "RNA : 2h -> 4h"); names(CPS.labs) <- c("dCPS_1and2", "dCPS_2and4")
cdf_common +
  stat_ecdf(data = sample2,
            aes(x = dCPS_value,
                col = dTED_group),
            geom = "step") +
  labs( title =expression("Txn DN: 0h"%->%"4h")) +facet_grid(.~dCPS_diff,labeller=labeller( dCPS_diff=CPS.labs))+labs( col =expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  xlab(expression(italic(Delta)*"RNA"))+labs(col=expression(italic(Delta)*"PAL"))#+
dev.off()

pal.labs <- c("PAL : DN", "PAL : NC"); names(pal.labs) <- c("DN", "NC")
col = yk.col(9)[c(2,8)] 
cdf_common = ggplot() +
  scale_color_manual(values = col) +
  theme_bw() +
  xlim(-2,2) +
  ylab("Cumulative fraction")
pdf("F5B.txnDN.CDFplot.ver2.pdf",width=4.5,height=2.5)
cdf_common +
  stat_ecdf(data = sample2,
            aes(x = dCPS_value,
                col = dCPS_diff),
            geom = "step") +
  labs( title =expression("Txn DN: 0h"%->%"4h")) +facet_grid(.~dTED_group,labeller=labeller( dTED_group=pal.labs))+labs( col =expression(italic(Delta)*"PAL (1h"%->%"2h)"))+
  xlab(expression(italic(Delta)*"RNA"))+labs(col=expression(italic(Delta)*"RNA"))+ scale_color_manual( 
                        labels = c(expression("1h" %->%"2h"), expression("2h" %->% "4h")),values=col)
dev.off()

```

Sampling for FigS4

Transcription downregulation genes.
```{R}
tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0<(-1)&PRO_1-PRO_0<=0&PRO_2-PRO_0<=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("PRO"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)
#Validaion no difference in PRO-seq profiles. among TED groups.
sample_=sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
  mutate(.PRO_0=0)%>%mutate(.PRO_1=PRO_1-PRO_0)%>%
  mutate(.PRO_2=PRO_2-PRO_0)%>%mutate(.PRO_4=PRO_4-PRO_0)%>%
  gather("time","txn",starts_with(".PRO_"))%>%
  separate(time,c("PRO","hour"),sep="_")%>%mutate(hour=as.numeric(hour))%>%
  select(dTED_group,hour,txn,newid)
df_summary=sample_%>%group_by(dTED_group,hour)%>%summarise(n=n(),mean=mean(txn),sd=sd(txn),se = sd(txn, na.rm=T)/sqrt(sum(!is.na(txn))))
df_summary$CI_lower <- df_summary$mean + qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
df_summary$CI_upper <- df_summary$mean - qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
#df_summary$CI_lower <- df_summary$mean + df_summary$sd
#df_summary$CI_upper <- df_summary$mean - df_summary$sd
col = yk.col(9)[c(2,5)] 
#col = yk.col(10)[c(3,5,7)] 
pdf("F5A.txndn.pdf",width=4,height=2.3)
pal.labs <- c("PAL : DN", "PAL : NC", "PAL : UP"); names(pal.labs) <- c("DN", "NC","UP")
ggplot(df_summary, aes(x=hour, y=mean)) +
   geom_line(data=df_summary, aes(x=hour, y=mean,group=dTED_group), size=1, alpha=0.8)+geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper,group=dTED_group,fill=dTED_group), alpha=0.3)+scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+theme_bw()+
  facet_grid(.~dTED_group,labeller = labeller(dTED_group=pal.labs))+
  #scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+
  ylab("mean(Transcription) ± 95% CI")
dev.off()
```
#  Time-course CPS-seq profiles. among TED groups.

```{r}

tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0<(-1)&PRO_1-PRO_0<=0&PRO_2-PRO_0<=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("CPS_"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)


sample_=sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO)%>%
mutate(.CPS_0=0)%>%mutate(.CPS_1=log2((CPS_1+1)/(CPS_0+1)))%>%mutate(.CPS_2=log2((CPS_2+1)/(CPS_0+1)))%>%mutate(.CPS_4=log2((CPS_4+1)/(CPS_0+1)))%>%
  gather("time","RNA",starts_with(".CPS_"))%>%separate(time,c("cps","hour"),sep="_")%>%mutate(hour=as.numeric(hour))%>%select(dTED_group,hour,RNA,newid)
df_summary=sample_%>%group_by(dTED_group,hour)%>%summarise(n=n(),mean=mean(RNA),sd=sd(RNA),se = sd(RNA, na.rm=T)/sqrt(sum(!is.na(RNA))))
df_summary$CI_lower <- df_summary$mean + qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
df_summary$CI_upper <- df_summary$mean - qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
#df_summary$CI_lower <- df_summary$mean + df_summary$sd
#df_summary$CI_upper <- df_summary$mean - df_summary$sd


pdf("F5C.TxnDN.pdf",width=4,height=2.3)
pal.labs <- c("PAL : DN", "PAL : NC", "PAL : UP"); names(pal.labs) <- c("DN", "NC","UP")

ggplot(df_summary, aes(x=hour, y=mean)) +
   geom_line(data=df_summary, aes(x=hour, y=mean,group=dTED_group), size=1, alpha=0.8)+geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper,group=dTED_group,fill=dTED_group), alpha=0.3)+theme_bw()+
  facet_grid(.~dTED_group, labeller = labeller(dTED_group=pal.labs))+
  scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+ylab("mean(RNA) ± 95% CI")+ylim(c(-1.5,1))

dev.off()

```

PAL time-course
```{R}


tmp=tmp=final%>%
  #transcription up.
  filter(PRO_4-PRO_0<(-1)&PRO_1-PRO_0<=0&PRO_2-PRO_1<=0)%>%
  mutate(dPRO=PRO_4-PRO_0)%>%
  mutate("dCPS_1and2"=log2((CPS_2+1)/(CPS_1+1)))%>%
  mutate("dCPS_2and4"=log2((CPS_4+1)/(CPS_2+1)))%>%
  #mutate("dCPS_value"=log2((CPS_1+1)/(CPS_0+1)))%>%
  mutate("dTED_value"=TED_2-TED_1)%>%
  select(newid,newid_old,starts_with("d"),contains("TED_"))%>%
  mutate("dTED_group"=ifelse(dTED_value<(-6),"DN",ifelse(dTED_value>6,"UP",ifelse(abs(dTED_value)<3,"NC","NA"))))%>%filter(dTED_group!="NA"&dTED_group!="UP")
sampln_=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%summarise(sum=n())%>%ungroup()%>%
  complete(dTED_group,nesting(dPRO_bin),fill =list(sum=0))%>%group_by(dPRO_bin)%>%summarise(sum=min(sum))

nested_sample=tmp%>%
  mutate(dPRO_bin=cut(dPRO,breaks=dpro.breaks))%>%
  group_by(dTED_group,dPRO_bin)%>%
  nest()%>%
  ungroup()%>%
  inner_join(sampln_,by="dPRO_bin")%>%ungroup()

sampled<-nested_sample%>%
  mutate(samp=map2(data,sum,sample_n))
sample=sampled%>%select(-data)%>%unnest(samp)


sample_=sample%>%select(-dPRO_bin,-sum,-newid_old,-dPRO,-dTED_value,-dCPS_1and2,-dCPS_2and4)%>%
mutate(.TED_0=0)%>%
  mutate(.TED_1=TED_1-TED_0)%>%
  mutate(.TED_2=TED_2-TED_0)%>%
  mutate(.TED_4=TED_4-TED_0)%>%
  gather("time","PAL",starts_with(".TED_"))%>%separate(time,c("ted","hour"),sep="_")%>%mutate(hour=as.numeric(hour))%>%select(dTED_group,hour,PAL,newid)
df_summary=sample_%>%group_by(dTED_group,hour)%>%summarise(n=n(),mean=mean(PAL),sd=sd(PAL),se = sd(PAL, na.rm=T)/sqrt(sum(!is.na(PAL))))
df_summary$CI_lower <- df_summary$mean + qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
df_summary$CI_upper <- df_summary$mean - qt((1-0.95)/2, df=df_summary$n-1)*df_summary$se
#df_summary$CI_lower <- df_summary$mean + df_summary$sd
#df_summary$CI_upper <- df_summary$mean - df_summary$sd


pdf("F5C.TXNDN.PALtimecourse.pdf",width=4,height=2.3)
pal.labs <- c("PAL : DN", "PAL : NC", "PAL : UP"); names(pal.labs) <- c("DN", "NC","UP")

ggplot(df_summary, aes(x=hour, y=mean)) +
   geom_line(data=df_summary, aes(x=hour, y=mean,group=dTED_group), size=1, alpha=0.8)+geom_ribbon(aes(ymin=CI_lower, ymax=CI_upper,group=dTED_group,fill=dTED_group), alpha=0.3)+theme_bw()+facet_grid(.~dTED_group, labeller = labeller(dTED_group=pal.labs))+scale_fill_manual(values=col,name = "PAL", labels = c("DN", "NC", "UP"))+ylab("mean(PAL) ± 95% CI")

dev.off()
```













0 vs 1 h

5. Select Genes with transcription upregulation.
```{R}
sample=final%>%select(c(starts_with("PRO"),starts_with("CPS"),starts_with("TED"),id))%>%
  mutate("dPRO_0and1"=PRO_1-PRO_0)%>%
  mutate("dCPS_0and1"=log2(CPS_1/CPS_0))%>%
  mutate("dTED_0and1"=TED_1-TED_0)%>%
  select(id,starts_with("d"))%>%
  mutate(dPROgroup=ifelse(dPRO_0and1>=1,"UP",ifelse(dPRO_0and1<(-1),"DN","NC")))%>%
  mutate(dPROgroup=factor(dPROgroup,levels=c("DN","NC","UP")))%>%
  #mutate("dCPS_0and1"=log2(CPS_1/CPS_0))%>%
  #mutate("dCPS_0and2"=log2(CPS_2/CPS_0))%>%mutate("dCPS_0and4"=log2(CPS_4/CPS_0))%>%
  filter(dPROgroup=="UP")%>%
  mutate("dTED_group"=ifelse(dTED_0and1<(-10),"PAL_DN",ifelse(dTED_0and1>(10),"PAL_UP",ifelse(abs(dTED_0and1)<5,"PAL_NC","NA"))))%>%
  filter(dTED_group!="NA")%>%
  gather("dCPS_diff","dCPS_value",starts_with("dCPS"))%>%filter(dCPS_diff=="dCPS_0and1")%>%
  #mutate("dCPS_group"=ifelse(dCPS_value<(-0.5),"DN",ifelse(dCPS_value>0.5,"UP","NC")))%>%
  group_by(dTED_group)%>%
  #sample_n(19)%>%
  ungroup()
col = yk.col(9)[c(2,5,8)] 


pdf("F5B.0.1.pdf",width=4.5,height=2.5)
ggplot(sample,aes(x=dTED_group,y=dCPS_value))+
  geom_quasirandom(aes(col=dTED_group),size=1)+
  scale_colour_manual(values=col )+theme_bw()+ylab(expression(italic(Delta)*"RNA(log2 FC (1h/0h))"))+
  geom_boxplot(alpha=0.3, color="black", width=.2)+xlab(expression(italic(Delta)*"PAL (0 h vs 1 h)"))+
  ggtitle("Genes with transcription up")
#+geom_beeswarm(size=2,aes(col=dTED_group))
dev.off()

```



