---
title: "Untitled"
author: "Yeonui+Kwak"
date: "7/23/2021"
output: html_document
---
```{r}
getwd()
library(RColorBrewer)
#nstall.packages("raster")
library(raster)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }

library(gridExtra)
#source("scatterPlot.R")
```

Load data

```{R}

dat=read.table("n2.genes.readcount.acrosstissues.txt",header=T,stringsAsFactors = F)
dat2=read.table("rc.per.transcript.hg19.bed",header=F,stringsAsFactors = F)
head(dat2)
dat2=dat2%>%select(V4,V8,V10,V12,V13)
colnames(dat2)=c("hg19.ENST","Exonid","gene","readcount.a375","blocksize")
dat2=dat2%>%group_by(hg19.ENST)%>%summarise(rc.a375=(sum(readcount.a375)*1000)/sum(blocksize))%>%ungroup()
dat=dat%>%inner_join(dat2,by="hg19.ENST")
#a375 readcount >5
dat%>%filter(coding=="protein_coding")%>%summarise(n_distinct(hg19.ENST))
dat%>%filter(rc.a375>1)%>%summarise(n_distinct(hg19.ENST))
#dat=dat%>%mutate(mean.bcell=(bcells.1+bcells.2)/2)
dat.long=dat%>%gather(-rc.a375,-chr,-start,-coding,-end,-id,-sumrc,-strand,-gene,-hg19.ENST,-CPSpos,-CPSposIndex,-pos,-n,key="tissue",value="rc")
geneid=dat%>%select(hg19.ENST,gene)%>%distinct_at(vars(hg19.ENST,gene))
(dat)
#4253genes

```

```{R}

dat=read.table("n2.genes.readcount.acrosstissues.txt",header=T,stringsAsFactors = F)
nrow(dat)/2
dat2=read.csv("hg19.rc.RPKM.threshold0.txt",header=T,stringsAsFactors = F,sep="\t")
nrow(dat2)
colnames(dat2)=c("gene","hg19.ENST","length","raw.readcount.a375","norm.count","RPKM")
dat2%>%filter(gene=="MIA")
dat2%>%
  filter(RPKM>1)%>%
  summarise(n_distinct(gene))


x=dat2%>%
  filter(RPKM>1)%>%select(gene)%>%unique()%>%unlist()
dat2%>%filter(!(gene %in% x))
dat=dat%>%inner_join(dat2%>%select(-gene),by="hg19.ENST")
#a375 readcount >5
dat%>%filter(coding=="protein_coding")%>%summarise(n_distinct(hg19.ENST))

dat%>%filter(RPKM>1)%>%summarise(n_distinct(hg19.ENST))
#dat=dat%>%mutate(mean.bcell=(bcells.1+bcells.2)/2)
dat.long=dat%>%gather(-raw.readcount.a375,-RPKM,-norm.count,-chr,-start,-coding,-end,-id,-sumrc,-strand,-gene,-hg19.ENST,-CPSpos,-CPSposIndex,-pos,-n,key="tissue",value="rc")
geneid=dat%>%select(hg19.ENST,gene)%>%distinct_at(vars(hg19.ENST,gene))
dat.long
#4253genes
56199-28389
```

```{R}

dis=read.table("PAS.distance.ratio.n2.txt",header=T,stringsAsFactors = F) #include miRNA, linc, protein coding

head(dis)

dis%>%filter(hg19.ENST=="ENST00000263741")
dis=dis%>%select(hg19.ENST,pos,distance1.2)%>%arrange(hg19.ENST)%>%distinct
dis
dat=dat%>%ungroup()%>%inner_join(dis%>%ungroup,by="hg19.ENST")
head(dat)



fin.bin=dat%>%mutate(bin=ifelse(distance1.2<100,"<100",ifelse(distance1.2<300,"100-300",ifelse(distance1.2<1000,"300-1000",ifelse(distance1.2<3000,"1000-3000",">3000")))))%>%mutate(bin=factor(bin,level=c("<100","100-300","300-1000","1000-3000",">3000")))%>%mutate(candidate=ifelse(distance1.2<300,"No candidate","Candiate"))




fin.bin2=fin.bin%>%distinct_at(vars(hg19.ENST),.keep_all=T)
dat%>%group_by(coding)%>%summarise(n_distinct(hg19.ENST))
fin.bin%>% filter(coding=="protein_coding")%>%group_by(bin)%>%summarise(n_distinct(hg19.ENST))

write.table(fin.bin%>%filter(bin %in% c("300-1000","1000-3000",">3000")),"genes with distance gt 300.txt",col.names = T,row.names = F,sep="\t",quote = F)



pdf("Genes with two PASs_distance.pdf",width=8,height=3)
ggplot(fin.bin2,aes(x=bin,fill=candidate))+geom_bar(stat="count")+facet_wrap(.~coding,scale="free")+theme_bw()+xlab("distance between PAS(proximal-distal,nt)")+
  ylab("number of genes")+ggtitle("Genes with two PASs")+scale_fill_manual(values=c("red","grey"))+
  theme(axis.text.x = element_text(angle = 45,hjust=1))#+xlim(c(-40,10000))
dev.off()


fin.bin%>%filter(RPKM>1)%>%group_by(coding,bin)%>%summarise(n_distinct(hg19.ENST))

fin.bin%>%
  filter(RPKM>1)%>%
  filter(distance1.2>=300)%>%group_by(coding)%>%summarise(n_distinct(hg19.ENST))
1423+1192+392
1395+1175+387
3238-2529
index=dat%>%select(id,CPSposIndex)
a375.expressed=fin.bin%>%
  filter(RPKM>1)%>%select(hg19.ENST)%>%unlist()
```

proximal vs distal using rna-seq readcounts in A375 mapped tothe entire genes

```{R}
utr=read.table("hg19.3UTR.1kbextension.bed11",header=T,stringsAsFactors = F)
head(utr)
utr=utr%>%select(chr,start,end,ENST,no,strand,Gene,lastexon)%>%rename(hg19.ENST=ENST)

dis=read.table("PAS.distance.ratio.n2.txt",header=T,stringsAsFactors = F) #include miRNA, linc, protein coding
dis%>%filter(hg19.ENST=="ENST00000263741")
dis=dis%>%select(hg19.ENST,pos,distance1.2)%>%arrange(hg19.ENST)%>%distinct
```

```{r}
utr.pl=utr%>%inner_join(dis,by="hg19.ENST")%>%group_by(hg19.ENST)%>%arrange(hg19.ENST,no)%>%separate_rows(pos,sep=",")%>%
  separate(pos,c("chr_","cord"),sep=":",remove=F)%>%ungroup()%>%mutate(cord=as.numeric(cord))%>%select(-chr_)%>%group_by(hg19.ENST,pos)%>%arrange(no)%>%filter(strand=="+")

utr.pl%>%group_by(hg19.ENST,pos)%>%summarise(n=n())%>%ungroup()%>%group_by(n)%>%summarise(n())

#ALL pas in a single 3UTR block
utr.pl=utr.pl%>%inner_join(utr.pl%>%group_by(hg19.ENST,pos)%>%summarise(n=n())%>%ungroup()%>%group_by(n)%>%filter(n==1)%>%distinct(hg19.ENST)%>%select(hg19.ENST),by="hg19.ENST")
utr.pl.tmp=utr.pl%>%arrange(hg19.ENST,pos)%>%group_by(hg19.ENST,pos)%>%mutate(newend=cord)%>%mutate(distance.stopcodon=cord-start)
utr.pl.tmp%>%filter(distance.stopcodon>150)

utr.mn=utr%>%inner_join(dis,by="hg19.ENST")%>%group_by(hg19.ENST)%>%arrange(hg19.ENST,no)%>%separate_rows(pos,sep=",")%>%
  separate(pos,c("chr_","cord"),sep=":",remove=F)%>%ungroup()%>%mutate(cord=as.numeric(cord))%>%select(-chr_)%>%group_by(hg19.ENST,pos)%>%arrange(no)%>%
  filter(strand=="-")
utr.mn%>%group_by(hg19.ENST,pos)%>%summarise(n=n())%>%ungroup()%>%group_by(n)%>%summarise(n())
utr.mn=utr.mn%>%inner_join(utr.mn%>%group_by(hg19.ENST,pos)%>%summarise(n=n())%>%ungroup()%>%group_by(n)%>%filter(n==1)%>%distinct(hg19.ENST)%>%select(hg19.ENST),by="hg19.ENST")
utr.mn.tmp=utr.mn%>%arrange(hg19.ENST,pos)%>%group_by(hg19.ENST,pos)%>%mutate(newstart=cord)%>%mutate(distance.stopcodon=end-cord)
utr.mn.tmp%>%filter(distance.stopcodon>150)

utr.pl.tmp=utr.pl.tmp%>%mutate(end=newend)%>%select(-newend)
utr.mn.tmp=utr.mn.tmp%>%mutate(start=newstart)%>%select(-newstart)
final.tmp=bind_rows(utr.pl.tmp%>%ungroup(),utr.mn.tmp%>%ungroup())
final=final.tmp%>%filter(distance.stopcodon>150)%>%group_by(hg19.ENST)%>%summarise(n=n_distinct(pos))%>%filter(n==2)%>%select(hg19.ENST)%>%ungroup()%>%inner_join(final.tmp,by="hg19.ENST")
final=final%>%unite("id",c(hg19.ENST,pos),sep="_")
head(final)
#utr%>%inner_join(dis,by="hg19.ENST")%>%group_by(hg19.ENST)%>%arrange(hg19.ENST,no)%>%separate_rows(pos,sep=",")%>%
  separate(pos,c("chr_","cord"),sep=":",remove=F)%>%ungroup()%>%mutate(cord=as.numeric(cord))%>%select(-chr_)%>%group_by(hg19.ENST,pos)%>%arrange(no)%>%summarise(n=n())%>%ungroup()%>%
  group_by(n)%>%filter(n==1)%>%summarise(n_distinct(hg19.ENST))

#final%>%summarise(n_distinct(hg19.ENST))
#final.tmp%>%summarise(n_distinct(hg19.ENST))
finalutr=final%>%select(id,chr,start,end,strand,Gene,distance1.2,n,distance.stopcodon)%>%rename(blockcount=n)%>%inner_join(index,by="id")

(finalutr)
write.table(finalutr,"hg19.3UTR.tandemPAS2.blockcount1.txt",col.names = T,row.names = F,sep="\t",quote = F)
```
#Save proximal PAS upsstream 150 nt into a bed file
```{R}
tmp=finalutr%>%filter(CPSposIndex==1)
tmp
tmp.mn=tmp%>%filter(strand=="-")%>%mutate(end=start+151)%>%select(chr,start,end,id,Gene,strand,distance1.2,distance.stopcodon)
tmp.pl=tmp%>%filter(strand=="+")%>%mutate(start=end-151)%>%select(chr,start,end,id,Gene,strand,distance1.2,distance.stopcodon)
tmp.final.upstream=rbind(tmp.mn,tmp.pl)
write.table(tmp.final.upstream,"pPAS.upstream150ntinterval.bed",col.names = F,row.names = F,sep = "\t",quote = F)

tmp.final.upstream%>%filter(distance1.2>0)
```

#Save proximal PAS DOWNstream 150 nt into a bed file

```{R}
tmp=finalutr%>%filter(CPSposIndex==1)
tmp
tmp.mn=tmp%>%filter(strand=="-")%>%mutate(end=start)%>%mutate(start=end-151)%>%select(chr,start,end,id,Gene,strand,distance1.2,distance.stopcodon)
tmp.pl=tmp%>%filter(strand=="+")%>%mutate(start=end)%>%mutate(end=start+151)%>%select(chr,start,end,id,Gene,strand,distance1.2,distance.stopcodon)
tmp.final.dnstream=rbind(tmp.mn,tmp.pl)
tmp.final.dnstream%>%filter(strand=="+")
write.table(tmp.final.dnstream,"pPAS.dnstream150ntinterval.bed",col.names = F,row.names = F,sep = "\t",quote = F)

```

#join the gRNA table to the UTR info table.
##canonical PAS signal reference
###Genome Res. 2000 Jul; 10(7): 1001–1010.  doi: 10.1101/gr.10.7.1001
```{r}
dat=read.table("hg19.3UTR.tandemPAS2.blockcount1.txt",header=T, stringsAsFactors = F)
head(dat)
dat.proximal=dat%>%filter(CPSposIndex==1)
nrow(dat.proximal)
head(dat.proximal)
```
```{R}
PASsignal=read.csv("../REsite/output.PASpos.upstream150interval.txt",header=T, stringsAsFactors = F,sep="\t")
PASsignal=PASsignal%>%gather("PAS.signal","position",-id)%>%mutate(position=substr(position,2,nchar(position)-1))%>%mutate(position=ifelse(position=="",NA,position))%>%spread(PAS.signal,position)
PASsignal2=PASsignal%>%gather("PAS.signal","position",-id)%>%group_by(id)%>%summarise(count.PASsignal=sum(!is.na(position)))%>%
  ungroup()%>%
  inner_join(PASsignal,by="id")
head(PASsignal2)
PASsignal2=PASsignal2%>%separate(id,c("id2","interval"),sep="::",remove = F)
```

```{R}
library(tidyr)
PASsignal3=PASsignal2%>%select(-id,-interval)%>%gather("PAS6mer","pos",-id2,-count.PASsignal)%>%separate_rows(pos,convert=T)%>%group_by(id2)%>%summarise(signal=any(pos>=100,na.rm=T))%>%ungroup()
PASsignal3%>%filter(signal==TRUE)%>%nrow()
PAS.indexid=PASsignal3%>%filter(signal==TRUE)%>%select(id2)%>%unlist()
  #%>%filter(id2=="ENST00000233969_chr2:103327118")
PASsignal2%>%select(-id,-interval)%>%gather("PAS6mer","pos",-id2,-count.PASsignal)%>%separate_rows(pos,convert=T)%>%group_by(id2)%>%filter(pos>=100)
```


```{R}
dat.prx.tmp=dat.proximal%>%rename("id2"=id)%>%inner_join(PASsignal2,by="id2")
head(dat.prx.tmp)
```

#upstream gRNA design
REsites.output.gRNA.upstream150interval.txt
```{r}
upgRNA=read.csv("../REsite/REsites.output.gRNA.upstream150interval.txt",header=T, stringsAsFactors = F,sep="\t")
upgRNA=upgRNA%>%mutate(region="upstream")
dngRNA=read.csv("../REsite/REsites.output.gRNA.dnstream150interval.txt",header=T, stringsAsFactors = F,sep="\t")
dngRNA=dngRNA%>%mutate(region="dnstream")

gRNA.all=rbind(upgRNA,dngRNA)
nrow(gRNA.all)
head(gRNA.all)

##upgRNA distance from stopcodon
gRNAindex.tmp=gRNA.all%>%separate(contig,c("id2","interval"),sep="::")%>%filter(region=="upstream")%>%inner_join(dat.proximal%>%rename("id2"=id),by="id2")
#gRNA vs Stopcodon distance>50
gRNAindex=gRNAindex.tmp%>%mutate(distance.gRNA.stopcodon=ifelse(strand=="+",distance.stopcodon-(150-start.x),distance.stopcodon-stop))%>%select(id2,target,distance.gRNA.stopcodon)%>%arrange(id2)%>%unite("newid",c(id2,target),sep="_")%>%filter(distance.gRNA.stopcodon>=50)%>%select(newid)%>%unlist()
head(gRNAindex2)

#gRNA vs PAS signal distance>50
gRNAindex2=gRNAindex.tmp%>%mutate(distance.gRNA.PASsignal=ifelse(strand=="+",(150-stop),start.x))%>%select(id2,target,distance.gRNA.PASsignal)%>%arrange(id2)%>%unite("newid",c(id2,target),sep="_")%>%filter(distance.gRNA.PASsignal>=50)%>%select(newid)%>%unlist()

```

##integrate PAS signal, 3UTR info, gRNA info
```{R}

dat.prx.final=dat.prx.tmp%>%inner_join(gRNA.all%>%separate(contig,c("id2","interval"),sep="::"),by="id2")%>%rename("gRNA.start"=start.y,"gRNA.stop"=stop)
nrow(dat.prx.final) #66861
colnames(dat.prx.final)
head(dat.prx.final)
```

##filtering
0. protein coding genes with two distinct PAS in a last 3'UTR exon (3'UTR composed of single exon).
1. distance (pPAS -stop codon >150 nt) #3071
2. pPAS should have conserved PAS signal upstream (-50 nt -0 nt) #2553
3. distance proximal PAS - distal PAS >=300 #2007
4. stopcodon-upsteam gRNA distance>=50 #
5. gRNA position upstream of pPAS PAS signal
6. gRNA poly(T) in down and upstream gRNAs X
7. gRNA restriction enzyme cut site in down and upstream gRNAs (x)

```{R}
dat.prx.final%>%summarise(n_distinct(id2)) #3071
dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%summarise(n_distinct(id2))#2553

dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%
  filter(distance1.2>=300)%>%
  summarise(n_distinct(id2)) #2007


#gRNA
dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%
  filter(distance1.2>=300)%>%group_by(count.PASsignal)%>%summarise(n_distinct(id2))
nrow(dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%
  filter(distance1.2>=300))

dn=dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%
  filter(distance1.2>=300)%>%
  unite("newid",c(id2,target),sep="_",remove=F)%>%filter(region=="dnstream")
#up=dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%filter(distance1.2>=300)%>% unite("newid",c(id2,target),sep="_",remove=F)%>%filter(region=="upstream")

up=dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%
  filter(distance1.2>=300)%>%
  unite("newid",c(id2,target),sep="_",remove=F)%>%filter(region=="upstream")%>%filter(newid %in% gRNAindex)
dat.prx.final2=rbind(dn,up)
nrow(dat.prx.final)
nrow(dn);nrow(up)
nrow(dn);nrow(up)
dn%>%mutate(sum.RE=BsmBI+NdeI+MluI)%>%filter(sum.RE==0)%>%filter(dangerous_polyT!="PolyT")%>%nrow()
dat.prx.final2%>%group_by(id2,region)%>%summarise(n.gRNA=n_distinct(target))%>%ungroup()%>%spread(region,n.gRNA)%>%filter(dnstream>4&upstream>4)#gRNA startpoint from stopcodon>50 #1984
dat.prx.final2%>%filter(id2=="ENST00000159111_chr19:5151901")
```


```{R}


up=dat.prx.final%>%filter(id2 %in% PAS.indexid)%>%
  filter(distance1.2>=300)%>%
  unite("newid",c(id2,target),sep="_",remove=F)%>%filter(region=="upstream")%>%filter(newid %in% gRNAindex)%>%filter(newid %in% gRNAindex2)
nrow(up)
up%>%mutate(sum.RE=BsmBI+NdeI+MluI)%>%filter(sum.RE==0)%>%
  filter(dangerous_polyT!="PolyT")%>%
  nrow()
dat.prx.final2=rbind(dn,up)




```


```{R}


#restriction enzyme sites
dat.prx.final2%>%mutate(sum.RE=BsmBI+NdeI+MluI)%>%filter(sum.RE==0)%>%
  filter(dangerous_polyT!="PolyT")%>%
  group_by(id2,region)%>%summarise(n.gRNA=n_distinct(target))%>%ungroup()%>%spread(region,n.gRNA)%>%filter(dnstream>0&upstream>0) #1944

dat.prx.final2%>%
  filter(dangerous_polyT!="PolyT")%>%
  mutate(sum.RE=BsmBI+NdeI+MluI)%>%filter(sum.RE==0)%>%group_by(id2,region)%>%summarise(n.gRNA=n_distinct(target))%>%ungroup()%>%spread(region,n.gRNA)%>%filter(dnstream>4&upstream>4) #1942   #gRNA >=5 :1104 ROWS;gRNA >=3 :1575 ROWS


head(a375.expressed)
dat.prx.final2%>%
  filter(dangerous_polyT!="PolyT")%>%
  mutate(sum.RE=BsmBI+NdeI+MluI)%>%filter(sum.RE==0)%>%
  group_by(id2,region)%>%summarise(n.gRNA=n_distinct(target))%>%ungroup()%>%spread(region,n.gRNA)%>%
  filter(dnstream>=1&upstream>=1& dnstream*upstream>=5)%>%#1943
  #filter(dnstream>4&upstream>4)%>%
  separate(id2,c("hg19.ENST","CPSpos"),sep="_")%>%filter(hg19.ENST %in%a375.expressed)
1575-1214


finalid2=dat.prx.final2 %>%filter(dangerous_polyT!="PolyT")%>%
  mutate(sum.RE=BsmBI+NdeI+MluI)%>%filter(sum.RE==0)%>%group_by(id2,region)%>%summarise(n.gRNA=n_distinct(target))%>%ungroup()%>%spread(region,n.gRNA)%>%filter(dnstream>=1&upstream>=1& dnstream*upstream>=5)%>%separate(id2,c("hg19.ENST","CPSpos"),sep="_",remove = F)%>%filter(hg19.ENST %in%a375.expressed)#1943

final_gRNA_set=dat.prx.final2%>%filter(id2 %in% c(finalid2$id2))


dis=read.table("PAS.distance.ratio.n2.txt",header=T,stringsAsFactors = F) #include miRNA, linc, protein coding
dis%>%filter(hg19.ENST=="ENST00000263741")
dis=dis%>%select(hg19.ENST,pos,distance1.2)%>%arrange(hg19.ENST)%>%distinct
final.list=final_gRNA_set%>%select(-newid,-id,-CPSposIndex,-interval.x,-interval.y)%>%mutate(stopcodon=ifelse(strand=="+",start.x,end))%>%select(-start.x,-end)
final.list=final.list%>%separate(id2,c("hg19.ENST","CPSpos"),sep="_")%>%select(-CPSpos)%>%inner_join(dis%>%select(-distance1.2),by="hg19.ENST")%>%arrange(Gene,hg19.ENST,region,-Hsu2013,-Doench2014OnTarget)
colnames(final.list)
final.list%>%summarise(n_distinct(hg19.ENST))
write.table(final.list%>%select(  "Gene"   ,  "hg19.ENST"   ,               "chr",           "pos"     ,          "strand" ,    "blockcount" ,               "stopcodon" ,          "distance.stopcodon"  ,                                     "distance1.2",                    "region"    ,           "target" ,                    "context"  , "overflow"     ,              "orientation"   ,             "Doench2014OnTarget" ,        "DoenchCFD_maxOT"  ,          "DoenchCFD_specificityscore",  "Hsu2013"  ,      
"dangerous_GC"      ,         "dangerous_polyT"   ,         "dangerous_in_genome"   ,               "basesDiffToClosestHit" ,    
"closestHitCount"  ,          "X0.1.2.3.4_mismatch"   ,     "otCount"                    ),"20210811.finalcandidate.PAS.n2.txt",sep="\t",col.names = T,row.names = F,quote = F)

```

```{R}
fin.bin
t=fin.bin%>%select(gene.x,id,hg19.ENST,pos.x,n,coding,RPKM,distance1.2,bin,candidate)%>%inner_join(final%>%select(strand,distance.stopcodon,id),by="id")
t 
t%>%ungroup()%>%filter(bin %in% c("300-1000","1000-3000",">3000"))%>%filter(RPKM<=1)%>%summarise(n_distinct(hg19.ENST))
t%>%ungroup()%>%filter(bin %in% c("300-1000","1000-3000",">3000"))%>%filter(RPKM>1)%>%summarise(n_distinct(hg19.ENST))

t=fin.bin%>%select(gene,hg19.ENST,pos.x,n,coding,rc.a375,distance1.2,bin,candidate)%>%inner_join(final.tmp%>%select(hg19.ENST,strand,distance.stopcodon)%>%distinct())
write.table(t,"final.list.of.genes for screening.txt",col.names = T,row.names = F,sep="\t",quote = F)
t%>%ungroup()%>%filter(bin %in% c("300-1000","1000-3000",">3000"))%>%filter(rc.a375>5)%>%filter(gene=="CDK13")


```
