---
title: "PieChart"
author: "Yeonui+Kwak"
date: "9/6/2020"
output: html_document
---

1.Set a working directory
```{R}
setwd("~/Desktop/Publication_Mar/Final_Fig1/Rcode/piechart")

```

2. Setup color and load packages.
```{r}
#install.packages("raster")
library(RColorBrewer)
library(raster)
library(tidyverse)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggbeeswarm)
library(gridExtra)
source("~/Desktop/Publication_Mar/Final_Supplement/Rsourcecode/scatterPlot.R")
```
```{R}
yk.col = function(n = 5, sat = 0.7, lum = 1) {
	col = c("royalblue4", "deepskyblue3", "turquoise3",
	   "grey62", "goldenrod2", "orange2", "orangered2") %>%
	col2rgb %>%
	rgb2hsv
	return(colorRampPalette(hsv(col[1,], col[2,] * sat,
				    col[3,] * lum))(n)) }
```



3. load the PAS positions with readcounts greater than 5.
n=47986

```{R}
cps=read.table("cps.all.PASpositions.internalpAremoved.gt5.bed", stringsAsFactors = F, header=F)
three.utr=read.table("final.30980cps.in.3UTR.1kb.ext.txt", stringsAsFactors = F, header=T)
three.utr %>%summarise(n_distinct(CPSid))
```

4. Remove the PASs in the 3'UTR
-17001 PASs are not in the 3'UTR
```{R}
cps%>%filter(!(V4 %in% three.utr$CPSid))%>%nrow() #17001
cps.1=cps%>%filter(!(V4 %in% three.utr$CPSid))
```
5. PASs in the intron
4474 are in intron
12527 are not in the intron or 3'UTR, either.
```{R}
intron=read.table("PASs.Intron.txt", stringsAsFactors = F, header=F)

cps.1%>%filter(V4 %in% intron$V1)%>%summarise(n_distinct(V4)) #4474
cps.1%>%filter(!(V4 %in% intron$V1))%>%nrow()
cps.2=cps.1%>%filter(!(V4 %in% intron$V1))
```

6. PASs in the Intron
8639 in CDS
3888 are not in CDS, 3UTR, or Introns.
```{r}
CDS=read.table("PASs.CDS.txt",stringsAsFactors = F, header=F)
cps.2%>%filter(V4 %in% CDS$V1)%>%summarise(n_distinct(V4)) #8639
cps.3=cps.2%>%filter(!(V4 %in% CDS$V1))
nrow(cps.3)
```


7. PASs in the 5'UTR
464 IN 5'UTR
3424 are in intergentic regions
```{r}
five.utr=read.table("PASs.5UTR.txt",stringsAsFactors = F, header=F)
cps.3%>%filter(V4 %in% five.utr$V1)%>%summarise(n_distinct(V4)) #464
cps.3%>%filter(!(V4 %in% five.utr$V1))%>%nrow()
```
8. Draw a pie chart to explain the distribution of experimentally determined PASs in the genome.

```{R}
pie=data.frame("location"=c("5'UTR","Intron","CDS","3'UTR","Intergenic"),"PASs"=c(464,4474,8639,30980,3424))
data <-  pie%>%
  arrange(desc(location))%>%
  mutate(prop = round(PASs / sum(pie$PASs) *100,1)) %>%
  mutate(ypos = cumsum(prop)- 0.5*prop )

mycols <- yk.col(5)[1:5]
pdf("Fig1_Piechart2.pdf",height=3,width=4)
ggplot(data, aes(x = "", y = prop, fill = location)) +
  geom_bar(width = 1, stat = "identity", color = "white") +
  coord_polar("y", start = 0)+
  geom_text(aes(y = ypos, label =paste0(prop)), color = "white")+
  scale_fill_manual(values = mycols) +
  theme_void()+ggtitle("Genomic position of PASs")

dev.off()


```